---
title: "Analysis of Fenton Hypotheses"
author: "James Mickley"
output:
  html_document:
    keep_md: yes
    theme: readable
  html_notebook:
    fig_caption: yes
    theme: readable
graphics: yes
---

```{r setup, include = F}
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ Knitr Options
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

# Set root directory to the project directory
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())


# Set default knitr options: 
# Suppress warnings and messages, cache chunks, 
#  set default figure size to 6x8 at 300 dpi, and save a png and pdf
knitr::opts_chunk$set(warning = F, message = F, collapse = T, cache = T,
    fig.height = 6, fig.width = 8, dpi = 300, # 6x8" @ 300dpi:1800x2400=4.3MP
    dev = c('png', 'pdf'), dev.args = list(pdf = list(onefile = F)))

```

### Overview

We set up a number of expectations for the Fenton site

1. The temperature should be lower in the woods
2. The humidity should be higher in the woods
3. The light levels should be lower in the woods
4. Soil moisture should show a gradient in both transects
    a. The woods should have a gradient from top to bottom of hill
    b. The meadow should have a gradient from the meadow to the swamp



```{r 'Main_Code', include = F}

# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ Setup - This code is run, but output is hidden
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

# Load Packages
require(cowplot) # Needed for publication-quality ggplots
require(tidyr) # Needed for data wrangling
require(lubridate) # Needed for date wrangling
require(knitr) # Needed to print tables to markdown
require(dplyr) # Needed for data wrangling, load last to avoid conflicts

# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ Data Preparation
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

# Import datasets
data <- read.csv("Data/EMU-6-21/fentondata-all.csv")

# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ ggPlot Theme
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

ggplot.theme <- theme(
    
    # Text size for axis ticks
    axis.text.y = element_text(size = 14),
    axis.text.x = element_text(size = 14),
    
    # Text size for axis labels
    # Also move them away from the axes a bit for more space
    axis.title.x = element_text(size = 18, face = "bold", vjust = -1),
    axis.title.y = element_text(size = 18, face = "bold", vjust = 1.5),
    
    # Plot title size
    plot.title = element_text(size = 20, face = "bold"),
    
    # Margins for top, right, bottom, left
    plot.margin = grid::unit(c(1.5, 1.5, 1.5, 1.2), "lines"), 
    
    # Legend text size
    legend.text = element_text(size = 14),
    legend.text.align = 0, 
    legend.title = element_text(size = 16, face = "bold"),
    legend.key.size = grid::unit(1.4, "line"),
    legend.key = element_blank()
    )


```


```{r 'Data_Wrangling', echo = F, include = F}


# Data wrangling on the main dataset
fenton <- data %>%
    
    # Convert the timestamp to a date/time object
    mutate(timestamp = as.POSIXct(timestamp)) %>%
    
    # Remove the seconds from the time, leaving only hour/minute
    mutate(timestamp = timestamp - second(timestamp))

# Take a look at the structure of the main dataset
str(fenton)


# Convert the dataset to wide format 
fenton.wide <- fenton %>%

    # Go to pure long format: all measurements in one column
    gather(measurement, value, temperature, humidity, vwc, 
        par, light, soil, pressure) %>%
    
    # Make a column that is both source & measurement type
    unite(measure, source, measurement) %>%
    
    # Get rid of columns that aren't needed
    # Or won't work with wide format
    select(-site, -voltage, -type, -emu) %>%
    
    # Filter out the rows where the value is NA 
    filter(!is.na(value)) %>%
    
    # Go from long format to wide: each row is a site at one time
    spread(measure, value)


# Take a look at the structure of the wide dataset
str(fenton.wide)



# Calculate hourly means and differences between 2 transects @ Fentonb
diff <- fenton %>%
    
    filter(source == "EMU") %>%

    # Group by transect and each hour separately
    group_by(hour, transect) %>%
    
    # Calculate mean and variance by hour for each transect
    mutate(t.mean = mean(temperature, na.rm = T), 
        t.var = var(temperature, na.rm = T), 
        t.obs = sum(!is.na(temperature)),
        h.mean = mean(humidity, na.rm = T), 
        h.var = var(humidity, na.rm = T), 
        h.obs = sum(!is.na(humidity)),
        l.mean = mean(par, na.rm = T), 
        l.var = mean(par, na.rm = T), 
        l.obs = sum(!is.na(par))) %>%
    
    # Group by hour alone
    group_by(hour) %>%
    
    # Calculate difference between fenton woods and meadow for each hour
    # Also calculate standard error of differences between means & 95% conf.int
    # http://onlinestatbook.com/2/sampling_distributions/samplingdist_diff_means.html
    # http://stattrek.com/estimation/difference-in-means.aspx?Tutorial=AP
    mutate(t.diff = t.mean - t.mean[transect == "Meadow"][1], 
        t.vn = t.var / t.obs,
        t.stderr = sqrt(t.vn + t.vn[transect == "Meadow"][1]), 
        t.conf = t.stderr * 1.96,
        
        h.diff = h.mean - h.mean[transect == "Meadow"][1], 
        h.vn = h.var / h.obs,
        h.stderr = sqrt(h.vn + h.vn[transect == "Meadow"][1]), 
        h.conf = h.stderr * 1.96,
        
        l.diff = l.mean - l.mean[transect == "Meadow"][1], 
        l.vn = l.var / l.obs,
        l.stderr = sqrt(l.vn + l.vn[transect == "Meadow"][1]), 
        l.conf = l.stderr * 1.96
        
        ) %>%
    
    # Group by only the variables needed and summarize
    group_by(transect, day, hour, day.hr, t.mean, t.diff, t.stderr, t.conf, 
        h.mean, h.diff, h.stderr, h.conf, 
        l.mean, l.diff, l.stderr, l.conf) %>%
    
    # Summarize by transect & hour
    summarize %>%
    
    # Convert back to normal data frame
    data.frame




# Create a new dataset with just diff & conf.int
diff.long <- diff %>% 
    
    # Only include the woods, since differences will be zero @ Meadow (reference pt)
    filter(transect == "Woods") %>%
    
    # Get rid of unneeded columns
    select(-t.mean, -h.mean, -l.mean, -t.stderr, -h.stderr, -l.stderr) %>%
    
    # Gather diff and conf columns into variable & value columns
    gather(variable, value, ends_with("diff"), ends_with("conf")) %>%
    
    # Split the variable column into measurement and variable
    separate(variable, into = c('measurement','var')) %>%
    
    # Spread the variable & value columns back into diff and conf cols
    spread(var, value) %>%
    
    # Recode t, h, and l, into their measurement names
    mutate(measurement = factor(recode(measurement, 
        "t" = "Temperature", "h" = "Humidity", "l" = "PAR"), 
        levels = c("Temperature", "Humidity", "PAR"))) 



# Make a mean volumetric water content by hour/emu dataset
soil.hr <- fenton %>%

    # Group by emu and hour
    group_by(transect, emu, order, day, hour, day.hr) %>%
    
    # Summarize means and standard errors for each emu by hour
    summarize(mean = mean(vwc), stderr = sd(vwc) / sqrt(n())) %>%
    
    # Convert to normal data frame
    data.frame




# Figure out hours for darkness
darkness <- data.frame(dusk = seq(from = 8, to = 528, by = 24), 
        dawn = seq(from = 17, to = 528, by = 24))

# Duplicate the darkness dataset with different ymin/ymax by variable
diff.darkness <- rbind(
    darkness %>% mutate(measurement = "Temperature", ymin = -14, ymax = 8), 
    darkness %>% mutate(measurement = "Humidity", ymin = -26, ymax = 37), 
    darkness %>% mutate(measurement = "PAR", ymin = -1400, ymax = 10)) %>%
    
    # Order the measurements
    mutate(measurement = factor(measurement, 
        levels = c("Temperature", "Humidity", "PAR")))

# Make a darkness dataset for soil moisture
soil.darkness <- darkness %>%
    mutate(ymin = 0, ymax = 0.45)

```

## Missing Data

### EMUs

```{r 'Missing_Data', echo = F}

# Calculate percentage of missing data by EMU
missing.data <- fenton %>%
    
    # Only use data from EMUs
    filter(source == "EMU") %>%
    
    # Select the variables to work with
    select(emu, temperature, humidity, par, vwc) %>%

    # Group by the EMUs
    group_by(emu) %>%
    
    # Sum up the number of logs, and calculate percent missing
    mutate(logs = n(), logs.missing = 1 - (logs/2109)) %>%
    
    # Ungroup
    ungroup %>%
    
    # Go to pure long format: all measurements in one column
    gather(measurement, value, temperature, humidity, vwc, 
        par) %>%
    
    # Group by emu/measurement
    group_by(emu, logs, logs.missing, measurement) %>%
    
    # Summarize the number of non-NA values
    summarize(n = sum(!is.na(value))) %>%
    
    # Group by EMU
    group_by(emu) %>%
    
    # Calculate the number of points missing
    mutate(data.missing = 1 - sum(n) / (2109*4)) %>%
    
    # Spread the measurements back out into columns
    spread(measurement, n) %>%
    
    # Select column order
    select(emu, logs, logs.missing, temperature, humidity, par, 
        vwc, data.missing) %>%
    
    
    # Ungroup  
    ungroup
    

# Add a line for totals
missing.data %>%

    # Total everything up
    summarize(emu = "Total", logs = sum(logs), 
        logs.missing = 1 - (logs/(2109*8)), 
        temperature = sum(temperature), 
        humidity = sum(humidity), 
        par = sum(par), vwc = sum(vwc), 
        data.missing = 1 - (temperature + humidity + par + vwc) / 
            (2109*4*8)
        ) %>%
    
    # Add to the summaries of each EMU
    rbind(missing.data, .) %>%
    
    # Roud missing data to 3 decimals
    mutate_each(funs(round(.,3)), logs.missing, data.missing) %>%
    
    # Print a markdown table
    kable(digits = 3, format = "markdown")


```


### iButtons



### Hobo Pendants



### Hobo Microstations


## Main Hypotheses


```{r 'Hourly_Differences', echo = F}

# Take the long dataset
diff.long %>%

    # Graph differerences by hour
    ggplot(aes(x = hour, y = diff, color = measurement)) + 

        # Facet by measurement
        facet_wrap(~ measurement, ncol = 1, scales = "free") + 
    
        # Add rectangles for darkness
        geom_rect(data = diff.darkness, 
            aes(xmin = dusk, xmax = dawn, ymin = ymin, ymax = ymax, 
                x = NULL, y = NULL, color = NULL), 
            fill = "grey95", show.legend = F) + 
        
        # Add the points, colored by    
        geom_point() + 
    
        # http://ggplot2.tidyverse.org/reference/geom_smooth.html
        # http://www.fromthebottomoftheheap.net/2014/05/09/modelling-seasonal-data-with-gam/
        #geom_smooth(aes(x = day.hr), method = "gam", 
        #    formula = y ~ s(x, bs = "cc", k = 12)) + 
    
        # Add a red line at 0
        geom_hline(yintercept = 0, color = "red") + 

        # Add 95% confidence intervals as errorbars
        #geom_errorbar(aes(ymin = diff - conf, ymax = diff + conf)) + 
    
        # Add axis labels and graph title
        ylab("Relative Difference") + xlab("Hour") + 
        ggtitle("Wood Difference Relative to Meadow") + 
    
        # Add the custom theme
        ggplot.theme


```




```{r 'Soil_Gradient', echo = F}

# Graph soil moisture gradients across transect
soil.hr %>%

    # Graph soil moisture by hour, coloring by order in the transect
    ggplot(aes(x = hour, y = mean, color = order, group = order)) + 

        # Facet by transect
        facet_wrap(~ transect, ncol = 1) + 
        
        # Add lines for each EMU  
        geom_line() + 
    
        # Add rectangles for darkness
        geom_rect(data = soil.darkness, 
            aes(xmin = dusk, xmax = dawn, ymin = ymin, ymax = ymax, 
                x = NULL, y = NULL, color = NULL, group = NULL), 
            fill = "grey95", show.legend = F) + 
    
        # Color from red to blue
        scale_color_gradient(name = "EMU Order\nDry to Wet", 
            low = "blue", high = "red", trans = 'reverse') + 

        # Add 95% confidence intervals as errorbars
        geom_errorbar(aes(ymin = mean - stderr, ymax = mean + stderr)) + 
    
        # Restrict to 0 - 0.4 VWC
        coord_cartesian(ylim = c(0, 0.45)) + 
    
        # Add axis labels and graph title
        ylab("Volumetric Water Content") + xlab("Hour") + 
        ggtitle("Transect Moisture Gradients") + 
    
        # Add the custom theme
        ggplot.theme

```


## Other Stuff

```{r 'Hourly_Means', echo = F}

# Plot hourly means for each transect and eacy variable
diff %>%
    
    # Selec the hourly mean columns
    select(transect, hour, t.mean, l.mean, h.mean) %>%
    
    # Gather the mean columns into long format
    gather(measurement, mean, t.mean, l.mean, h.mean) %>%
    
    # Recode t, h, and l, into their measurement names
    mutate(measurement = factor(recode(measurement, 
        "t.mean" = "Temp", "h.mean" = "Humidity", "l.mean" = "Light"), 
        levels = c("Temp", "Light", "Humidity"))) %>%

    # Plot means over time, coloring by transect
    ggplot(aes(x = hour, y = mean, color = transect)) + 
        
        # Facet by measurement
        facet_wrap(~ measurement, ncol = 1, scales = "free") + 
        
        # Add points
        geom_line() +
    
        # Add axis labels and graph title
        ylab("Mean") + xlab("Hour") + 
        ggtitle("Hourly Means") + 
    
        # Add the custom theme
        ggplot.theme
    

```

