---
title: "Fenton Paired T-tests"
author: "James Mickley"
output:
  html_document:
    keep_md: yes
    theme: readable
  html_notebook:
    theme: readable
graphics: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include = F}
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ Knitr Options
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

# Set root directory to the project directory
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())


# Set default knitr options: 
# Suppress warnings and messages, cache chunks, 
#  set default figure size to 6x8 at 300 dpi, and save a png and pdf
knitr::opts_chunk$set(warning = F, message = F, collapse = T, cache = T,
    fig.height = 6, fig.width = 8, dpi = 300, # 6x8" @ 300dpi:1800x2400=4.3MP
    dev = c('png', 'pdf'), dev.args = list(pdf = list(onefile = F)))

```



## Overview

Paired T-tests between the two transects.  These pair observations by timestamp and order within the transect.  Any times/orders without paired data are thrown out.  

The difference in means is relative to the Woods transect 

```{r 'Main_Code', include = F, cache = F}

# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ Setup - This code is run, but output is hidden
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

# Load Packages
require(tidyverse) # Needed for data wrangling
require(lubridate) # Needed for date wrangling

# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ Data Preparation
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

# Import datasets
data <- read.csv("Data/EMU-6-21/fentondata-all.csv")


```



```{r 'Data_Wrangling', echo = F, include = F}

# Data wrangling on the main dataset
fenton <- data %>%
    
    # Convert the timestamp to a date/time object
    mutate(timestamp = as.POSIXct(timestamp)) %>%
    
    # Remove the seconds from the time, leaving only hour/minute
    mutate(timestamp = timestamp - second(timestamp))
    

# Take a look at the structure of the main dataset
str(fenton)

```


### Temperature

The woods is 1.8 ºC cooler than the meadow

```{r 'Temperature', echo = F, comment = ""}

# Paired t-test for temperature, based on hourly transect means
#fenton %>%
#    filter(source == "EMU") %>%
#    group_by(transect, hour) %>%
#    summarize(temperature = mean(temperature, na.rm = T)) %>%
#    ungroup %>% data.frame %>% 
#    t.test(temperature ~ transect, paired = T, data = .)


# Paired t-test for temperature
fenton %>%
    
    # Restrict to EMU data
    filter(source == "EMU") %>%

    # Select columns to work with
    select(transect, order, timestamp, temperature) %>%
    
    # Spread, to make one column for each transect
    spread(transect, temperature) %>%
    
    # Remove rows that don't have data for both transects
    filter(!is.na(Meadow), !is.na(Woods)) %>%
    
    # Gather the data back into a transect and temperature column
    gather(transect, temperature, Meadow, Woods) %>%
    
    # Reorder factor levels for transect
    mutate(transect = factor(transect, levels = c("Woods", "Meadow"))) %>%
    
    # Perform paired t-test
    t.test(temperature ~ transect, paired = T, data = .)


```


### Humidity

The woods is 3.4% more humid than the meadow

```{r 'Humidity', echo = F, comment = ""}


# Paired t-test for humidity
fenton %>%
    
    # Restrict to EMU data
    filter(source == "EMU") %>%

    # Select columns to work with
    select(transect, order, timestamp, humidity) %>%
    
    # Spread, to make one column for each transect
    spread(transect, humidity) %>%
    
    # Remove rows that don't have data for both transects
    filter(!is.na(Meadow), !is.na(Woods)) %>%
    
    # Gather the data back into a transect and humidity column
    gather(transect, humidity, Meadow, Woods) %>%
    
    # Reorder factor levels for transect
    mutate(transect = factor(transect, levels = c("Woods", "Meadow"))) %>%
    
    # Perform paired t-test
    t.test(humidity ~ transect, paired = T, data = .)


```



### PFD

The woods gets less light than the meadow by about 306 µmol/m^2/s

```{r 'PAR', echo = F, comment = ""}


# Paired t-test for PFD
fenton %>%
    
    # Restrict to EMU data
    filter(source == "EMU") %>%

    # Select columns to work with
    select(transect, order, timestamp, pfd) %>%
    
    # Spread, to make one column for each transect
    spread(transect, pfd) %>%
    
    # Remove rows that don't have data for both transects
    filter(!is.na(Meadow), !is.na(Woods)) %>%
    
    # Gather the data back into a transect and par column
    gather(transect, pfd, Meadow, Woods) %>%
    
    # Reorder factor levels for transect
    mutate(transect = factor(transect, levels = c("Woods", "Meadow"))) %>%
    
    # Perform paired t-test
    t.test(pfd ~ transect, paired = T, data = .)


```



### VWC

The woods is drier than the meadow by about 0.13 m^3/m^3

```{r 'VWC', echo = F, comment = ""}


# Paired t-test for vwc
fenton %>%
    
    # Restrict to EMU data
    filter(source == "EMU") %>%

    # Select columns to work with
    select(transect, order, timestamp, vwc) %>%
    
    # Spread, to make one column for each transect
    spread(transect, vwc) %>%
    
    # Remove rows that don't have data for both transects
    filter(!is.na(Meadow), !is.na(Woods)) %>%
    
    # Gather the data back into a transect and vwc column
    gather(transect, vwc, Meadow, Woods) %>%
    
    # Reorder factor levels for transect
    mutate(transect = factor(transect, levels = c("Woods", "Meadow"))) %>%
    
    # Perform paired t-test
    t.test(vwc ~ transect, paired = T, data = .)


```


### Session Information

```{r 'Session_Info', echo = F, comment = ""}

# Add session information to help with reproduceability
sessionInfo()


```
