---
title: "Comparison Among Sensor Types"
author: "James Mickley"
output:
  html_document:
    keep_md: yes
    theme: readable
  html_notebook:
    theme: readable
graphics: yes
---

```{r setup, include = F}
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ Knitr Options
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

# Set root directory to the project directory
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

# Set default knitr options: 
# Suppress warnings and messages, cache chunks, 
#  set default figure size to 6x8 at 300 dpi, and save a png and pdf
knitr::opts_chunk$set(warning = F, message = F, collapse = T, cache = T,
    fig.height = 6, fig.width = 8, dpi = 300, # 6x8" @ 300dpi:1800x2400=4.3MP
    dev = c('png', 'pdf'), dev.args = list(pdf = list(onefile = F)))

```



### Overview

This analysis compares the results for the EMU, Hobo Pendant, Hobo Microstation, and iButton at the first site in each of the two transects.  


```{r 'Main_Code', include = F, cache = F}

# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ Setup - This code is run, but output is hidden
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

# Load Packages
require(cowplot) # Needed for publication-quality ggplots
require(ggpmisc) # Needed for lm equations on graphs
require(tidyr) # Needed for data wrangling
require(lubridate) # Needed for date wrangling
require(dplyr) # Needed for data wrangling, load last to avoid conflicts

# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ Data Preparation
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

# Import datasets
data <- read.csv("Data/EMU-6-21/fentondata-all.csv")

# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ ggPlot Theme
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

ggplot.theme <- theme(
    
    # Text size for axis ticks
    axis.text.y = element_text(size = 14),
    axis.text.x = element_text(size = 14),
    
    # Text size for axis labels
    # Also move them away from the axes a bit for more space
    axis.title.x = element_text(size = 18, face = "bold", vjust = -1),
    axis.title.y = element_text(size = 18, face = "bold", vjust = 1.5),
    
    # Plot title size
    plot.title = element_text(size = 20, face = "bold"),
    
    # Margins for top, right, bottom, left
    plot.margin = grid::unit(c(1.5, 1.5, 1.5, 1.2), "lines"), 
    
    # Legend text size
    legend.text = element_text(size = 14),
    legend.text.align = 0, 
    legend.title = element_text(size = 16, face = "bold"),
    legend.key.size = grid::unit(1.4, "line"),
    legend.key = element_blank()
    )


```


```{r 'Data_Wrangling', echo = F, include = F}


# Data wrangling on the main dataset
data <- data %>%
    
    # Convert the timestamp to a date/time object
    mutate(timestamp = as.POSIXct(timestamp)) %>%
    
    # Remove the seconds from the time, leaving only hour/minute
    mutate(timestamp = timestamp - second(timestamp))

# Take a look at the structure of the main dataset
str(data)


# Convert the dataset to wide format 
data.wide <- data %>%

    # Go to pure long format: all measurements in one column
    gather(measurement, value, temperature, humidity, vwc, 
        par, light, soil, pressure) %>%
    
    # Make a column that is both source & measurement type
    unite(measure, source, measurement) %>%
    
    # Get rid of columns that aren't needed
    # Or won't work with wide format
    select(-site, -voltage, -type, -emu) %>%
    
    # Filter out the rows where the value is NA 
    filter(!is.na(value)) %>%
    
    # Go from long format to wide: each row is a site at one time
    spread(measure, value)


# Take a look at the structure of the wide dataset
str(data.wide)



# Figure out hours for darkness
darkness <- data.frame(dusk = seq(from = 8, to = 528, by = 24), 
        dawn = seq(from = 17, to = 528, by = 24))

diff.darkness <- rbind(
    darkness %>% mutate(measurement = "Temperature", ymin = -14, ymax = 8), 
    darkness %>% mutate(measurement = "Humidity", ymin = -26, ymax = 37), 
    darkness %>% mutate(measurement = "PAR", ymin = -1400, ymax = 10)) %>%
    
    # Order the measurements
    mutate(measurement = factor(measurement, 
        levels = c("Temperature", "Humidity", "PAR")))

```

### Temperature

Differences in temperature between EMUs and Hobos at all positions

Hobos tend to read high, especially during the day.  Strongly suggests heat trapped from sun.

```{r 'Temp_Diff_Hobo', echo = F}

# Set ymin and ymax limits separately for the two transects
temp.hobo.darkness <- rbind(
    darkness %>% mutate(transect = "Meadow", ymin = -6, ymax = 10),
    darkness %>% mutate(transect = "Woods", ymin = -4, ymax = 4))

# Plot relative differences in temperature
data.wide %>%
    
    # Make columns for differences between Hobo & EMU
    mutate(Hobo_diff = Hobo_temperature - EMU_temperature) %>%
    
    # Get rid of missing data rows
    filter(!is.na(Hobo_diff)) %>%
    
    # Plot differences over time
    ggplot(aes(x = hour, y = Hobo_diff)) +
    
        # Make a facet for each transect
        facet_wrap(~ transect, ncol = 1, scales = "free_y") + 
    
        # Add rectangles for darkness
        geom_rect(data = temp.hobo.darkness, aes(xmin = dusk, xmax = dawn, 
            ymin = ymin, ymax = ymax, x = NULL, y = NULL, color = NULL), 
            fill = "grey85", show.legend = F) + 
    
        # Add points
        geom_point(alpha = 0.3) +
    
        # Add horizontal line at zero
        geom_hline(yintercept = 0, color = "red") + 
    
        # Add axis labels
        xlab("Hour") + 
        ylab("Temperature Difference (ºC)") + 
    
        # Title
        ggtitle("Hobo Temperature Difference") +
    
        # Add theme
        ggplot.theme

```

Differences between EMUs and iButtons at the first position of each transect.

Prior to hour 98, the iButtons had fallen on the ground, so you are seeing the difference between ground temperature and that of 10 cm above.

There is minimal difference in temperature between the iButtons and Hobos on both transects from hour 98 to 190.  After 190, the meadow BME280 started failing, and I think those values may be bad.  

```{r 'Temp_Diff_iButton', echo = F}

# Set ymin and ymax limits separately for the two transects
temp.iButton.darkness <- rbind(
    darkness %>% mutate(transect = "Meadow", ymin = -8, ymax = 16),
    darkness %>% mutate(transect = "Woods", ymin = -8, ymax = 5))

# Plot relative differences in temperature
data.wide %>%
    
    # Make a colummn for differences between iButton & EMU
    mutate(iButton_diff = iButton_temperature - EMU_temperature) %>%

    # Get rid of missing data rows
    filter(!is.na(iButton_diff)) %>%
    
    # Plot differences over time
    ggplot(aes(x = hour, y = iButton_diff)) +
    
        # Make a facet for each transect
        facet_wrap(~ transect, ncol = 1, scales = "free_y") + 
    
        # Add rectangles for darkness
        geom_rect(data = temp.iButton.darkness, aes(xmin = dusk, xmax = dawn, 
            ymin = ymin, ymax = ymax, x = NULL, y = NULL, color = NULL), 
            fill = "grey85", show.legend = F) + 
    
        # Add points
        geom_point(alpha = 0.3) +
    
        # Add horizontal line at zero
        geom_hline(yintercept = 0, color = "red") + 

        # Add axis labels
        xlab("Hour") + 
        ylab("Temperature Difference (ºC)") + 
    
        # Title
        ggtitle("iButton Temperature Difference") +
    
        # Add theme
        ggplot.theme

```


### Humidity

Differences in humidity between EMUs and iButtons at the first position of each transect.

Prior to hour 98, the iButtons had fallen on the ground, so you are seeing the difference between ground temperature and that of 10 cm above.

It looks like humidity is lower during the day with the iButton on the meadow transect, and generally higher at all times in the woods.


```{r 'Humid_Diff', echo = F}

# Set ymin and ymax limits separately for the two transects
humid.darkness <- rbind(
    darkness %>% mutate(transect = "Meadow", ymin = -36, ymax = 55),
    darkness %>% mutate(transect = "Woods", ymin = -12, ymax = 72))


# Plot relative differences in humidity
data.wide %>%
    
    # Only include data from the first site on a transect
    #filter(order == 1, hour > 98) %>%
    
    # Make a colummn for differences between iButton & EMU
    mutate(iButton_diff = iButton_humidity - EMU_humidity) %>%

    # Get rid of missing data rows
    filter(!is.na(iButton_diff)) %>%
    
    # Plot differences over time
    ggplot(aes(x = hour, y = iButton_diff)) +
    
        # Make a facet for each transect
        facet_wrap(~ transect, ncol = 1, scales = "free_y") + 

        # Add rectangles for darkness
        geom_rect(data = humid.darkness, aes(xmin = dusk, xmax = dawn, 
            ymin = ymin, ymax = ymax, x = NULL, y = NULL, color = NULL), 
            fill = "grey85", show.legend = F) + 
    
        # Add points
        geom_point(alpha = 0.3) +
    
        # Add horizontal line at zero
        geom_hline(yintercept = 0, color = "red") + 
    
        # Theme the legend
        scale_color_manual(name = "Difference", values = c("blue")) + 
    
        # Add axis labels
        xlab("Date") + 
        ylab("Relative Humidity (%)") + 
    
        # Title
        ggtitle("iButton Humidity Difference") +
    
        # Add theme
        ggplot.theme

```


### PAR

Differences in PAR between EMUs and Hobo Microstations at the first position of each transect.

The Hobo is consistently lower in the woods, even at night (why is that?). There's a lot of noise in the meadow during the day.



```{r 'PAR_Diff', echo = F}

# Set ymin and ymax limits separately for the two transects
par.darkness <- rbind(
    darkness %>% mutate(transect = "Meadow", ymin = -500, ymax = 1650),
    darkness %>% mutate(transect = "Woods", ymin = -570, ymax = 320))

# Plot relative differences in PAR
data.wide %>%
    
    # Make a colummn for differences between iButton & EMU
    mutate(Hobo_diff = Hobo_par - EMU_par) %>%
    
    # Get rid of missing data rows
    filter(!is.na(Hobo_diff)) %>%
    
    # Plot differences over time
    ggplot(aes(x = hour, y = Hobo_diff)) +
    
        # Make a facet for each transect
        facet_wrap(~ transect, ncol = 1, scales = "free_y") + 

        # Add rectangles for darkness
        geom_rect(data = par.darkness, aes(xmin = dusk, xmax = dawn, 
            ymin = ymin, ymax = ymax, x = NULL, y = NULL, color = NULL), 
            fill = "grey85", show.legend = F) + 
    
        # Add points
        geom_point(alpha = 0.3) +
    
        # Add horizontal line at zero
        geom_hline(yintercept = 0, color = "red") + 
    
        # Add axis labels
        xlab("Hour") + 
        ylab(bquote(bold('PAR (µmol '~m^-2 ~s^-1~')'))) +
    
        # Title
        ggtitle("Hobo PAR Difference") +
    
        # Add theme
        ggplot.theme

```




### Volumetric Water Content

Differences in Volumetric Water Content between EMUs and Hobo Microstations at the first position of each transect.

The Hobo is always wetter in the woods, and it seems to be responding more strongly to rain than the EMU.

In the meadow, the pattern is unpredictable, but the measurements are closer together.

```{r 'VWC_Diff', echo = F}

# Set ymin and ymax limits separately for the two transects
vwc.darkness <- rbind(
    darkness %>% mutate(transect = "Meadow", ymin = -.08, ymax = 0.04),
    darkness %>% mutate(transect = "Woods", ymin = 0, ymax = 0.25))

# Plot relative differences in PAR
data.wide %>%
    
    # Make a colummn for differences between iButton & EMU
    mutate(Hobo_diff = Hobo_vwc - EMU_vwc) %>%
    
    # Get rid of missing data rows
    filter(!is.na(Hobo_diff)) %>%
    
    # Plot differences over time
    ggplot(aes(x = hour, y = Hobo_diff)) +
    
        # Make a facet for each transect
        facet_wrap(~ transect, ncol = 1, scales = "free_y") + 

        # Add rectangles for darkness
        geom_rect(data = vwc.darkness, aes(xmin = dusk, xmax = dawn, 
            ymin = ymin, ymax = ymax, x = NULL, y = NULL, color = NULL), 
            fill = "grey85", show.legend = F) + 
    
        # Add points
        geom_point(alpha = 0.3) +
    
        # Add horizontal line at zero
        geom_hline(yintercept = 0, color = "red") + 
    
        # Add axis labels
        xlab("Hour") + 
        ylab(bquote(bold('VWC'~(m^3/m^3)))) +

        # Title
        ggtitle("Hobo VWC Difference") +

        # Add theme
        ggplot.theme

```

## Comparisons between EMU & iButton/Hobo Pendant for Woods transect

The Hobo Pendant reads slightly high on temperature.  Presumably it's getting heated inside of its enclosed case.


```{r 'Temp_Comp_Hobo', echo = F, fig.width=10}


# Plot relative differences in temperature
data.wide %>%
    
    # Only show the woods transect at the first site
    filter(order == 1, !is.na(EMU_temperature)) %>%
    
    # Plot differences over time
    ggplot(aes(y = EMU_temperature, x = Hobo_temperature)) +
    
        # Make a facet for each transect
        facet_wrap(~ transect, ncol = 2) + 
    
        # Add points
        geom_point(color = "blue", alpha = 0.6) +
    
        # Add horizontal line at zero
        geom_abline(intercept = 0, slope = 1) + 
    
        # Add linear model
        geom_smooth(method = "lm", color = "red", se = F) + 
    
        # Add the equation
        stat_poly_eq(aes(label = ..eq.label..),
               eq.x.rhs = "italic(x)", coef.digits = 3, label.x = 15, 
               label.y = 40, parse = T, size = 7,
               formula = y ~ x) +
        
        # Add the R^2
        stat_poly_eq(aes(label = ..rr.label..),
               rr.digits = 3, label.x = 15, label.y = 36,
               parse = T, size = 7, formula = y ~ x) +
    
        # Add horizontal line at zero
        geom_abline(intercept = 0, slope = 1) + 
    
        # Add axis labels
        xlab("Hobo Temperature (ºC)") + 
        ylab("EMU Temperature (ºC)") + 
    
        # Add theme
        ggplot.theme

```

The iButtons read slightly lower than the EMU when it's cold, and slightly higher when hot.  Presumably, this is in response to the metal enclosure.

```{r 'Temp_Comp_iButton', echo = F, fig.width=10}

# Plot relative differences in temperature
data.wide %>%
    
    # Only show the woods transect at the first site
    filter(order == 1, hour > 98, 
        !is.na(EMU_temperature), !is.na(iButton_temperature)) %>%
    
    # Plot differences over time
    ggplot(aes(y = EMU_temperature, x = iButton_temperature)) +

        # Make a facet for each transect
        facet_wrap(~ transect, ncol = 2) + 
    
        # Add points
        geom_point(color = "blue", alpha = 0.6) +
    
        # Add horizontal line at zero
        geom_abline(intercept = 0, slope = 1) + 
    
        # Add linear model
        geom_smooth(method = "lm", color = "red", se = F) + 
    
        # Add the equation
        stat_poly_eq(aes(label = ..eq.label..),
               eq.x.rhs = "italic(x)", coef.digits = 4, label.x = 15, 
               label.y = 40, parse = T, size = 7,
               formula = y ~ x) +
        
        # Add the R^2
        stat_poly_eq(aes(label = ..rr.label..),
               rr.digits = 3, label.x = 15, label.y = 36,
               parse = T, size = 7, formula = y ~ x) +
    
        # Add horizontal line at zero
        geom_abline(intercept = 0, slope = 1) + 
    
        # Theme the legend
        #scale_color_manual(name = "Difference", values = c(2, 3)) + 
    
        # Add axis labels
        xlab("iButton Temperature (ºC)") + 
        ylab("EMU Temperature (ºC)") + 
    
        # Add theme
        ggplot.theme


```

### Humidity

The iButtons have slightly higher humidity (3-5% or so) in the Woods.  They also can return values higher than 100%.  

In the meadow, the BME280 failed, and I think all the humidity values are bogus.

```{r 'Humid_Comp', echo = F, fig.width=10}


# Plot relative differences in temperature
data.wide %>%

    # Only show the woods transect after hour 98 when iButtons  
    filter(order == 1, hour > 98, 
        !is.na(EMU_humidity), !is.na(iButton_humidity)) %>%
    
    # Plot differences over time
    ggplot(aes(x = iButton_humidity, y = EMU_humidity)) +

        # Make a facet for each transect
        facet_wrap(~ transect, ncol = 2) + 

        # Add points
        geom_point(color = "blue", alpha = 0.6) +
    
        # Add horizontal line at zero
        geom_abline(intercept = 0, slope = 1) + 

        # Add linear model
        geom_smooth(method = "lm", color = "red", se = F) + 
    
        # Add the equation
        stat_poly_eq(aes(label = ..eq.label..),
               eq.x.rhs = "italic(x)", coef.digits = 4, label.x = 50, 
               label.y = 100, parse = T, size = 7,
               formula = y ~ x) +
        
        # Add the R^2
        stat_poly_eq(aes(label = ..rr.label..),
               rr.digits = 3, label.x = 50, label.y = 95,
               parse = T, size = 7, formula = y ~ x) +
    
        # Theme the legend
        #scale_color_manual(name = "Difference", values = c(2, 3)) + 
    
        # Add axis labels
        xlab("iButton Humidity (%)") + 
        ylab("EMU Humidity (%)") + 
    
        # Add theme
        ggplot.theme


```


### PAR

Hobos tend to read too low in low light, and high in high light (sometimes wildly too high).

Roughly the same pattern in both transects.


```{r 'PAR_Comp', echo = F, fig.width=10}


# Plot relative differences in temperature
data.wide %>%

    # Only show the woods transect at the first site
    filter(order == 1, !is.na(EMU_par), !is.na(Hobo_par)) %>%
    
    # Plot differences over time
    ggplot(aes(x = Hobo_par, y = EMU_par)) +
    
        # Make a facet for each transect
        facet_wrap(~ transect, ncol = 2) + 

        # Add points
        geom_point(color = "blue", alpha = 0.6) +
    
        # Add horizontal line at zero
        geom_abline(intercept = 0, slope = 1) + 
    
        # Add linear model
        geom_smooth(method = "lm", color = "red", se = F) + 
    
        # Add the equation
        stat_poly_eq(aes(label = ..eq.label..),
               eq.x.rhs = "italic(x)", coef.digits = 4, label.x = 500, 
               label.y = 1500, parse = T, size = 7,
               formula = y ~ x) +
        
        # Add the R^2
        stat_poly_eq(aes(label = ..rr.label..),
               rr.digits = 3, label.x = 500, label.y = 1400,
               parse = T, size = 7, formula = y ~ x) +
    
        # Add axis labels
        xlab(bquote(bold('Hobo PAR (µmol '~m^-2 ~s^-1~')'))) +
        ylab(bquote(bold('EMU PAR (µmol '~m^-2 ~s^-1~')'))) +
    
        # Add theme
        ggplot.theme


```



### Volumetric Water Content

Very strange patterns here, and not at all what we saw during calibration.

In the meadow (which is wet), the hobo is mostly reading lower than the EMU, but sometimes it's higher.  While the trend is positive, the high readings on the Hobo are not at the higest readings for the EMU, but somewhere in the middle.

In the woods, the EMU read very low, while there was a full range for the Hobo.  
This was true of 3 of 4 EMUs along the transect.  Maybe it's a depth issue? Deeper soil, which the Hobo could access would be wetter.  Or perhaps it's a coarse soil issue?

```{r 'VWC_Comp', echo = F, fig.width=10}


# Plot relative differences in temperature
data.wide %>%

    # Only show the woods transect at the first site
    filter(order == 1, !is.na(EMU_par), !is.na(Hobo_par)) %>%
    
    # Plot differences over time
    ggplot(aes(x = Hobo_vwc, y = EMU_vwc)) +
    
        # Make a facet for each transect
        facet_wrap(~ transect, ncol = 2) + 

        # Add points
        geom_point(color = "blue", alpha = 0.6) +
    
        # Add horizontal line at zero
        geom_abline(intercept = 0, slope = 1) + 
    
        # Add linear model
        geom_smooth(method = "lm", color = "red", se = F) + 
    
        # Add the equation
        stat_poly_eq(aes(label = ..eq.label..),
               eq.x.rhs = "italic(x)", coef.digits = 4, label.x = 0.22, 
               label.y = 0.13, parse = T, size = 7,
               formula = y ~ x) +
        
        # Add the R^2
        stat_poly_eq(aes(label = ..rr.label..),
               rr.digits = 3, label.x = 0.22, label.y = 0.113,
               parse = T, size = 7, formula = y ~ x) +

        #xlim(0, 0.3) + ylim(0, 0.08) + 
    
        # Add axis labels
        xlab(bquote(bold('Hobo VWC'~(m^3/m^3)))) +
        ylab(bquote(bold('EMU VWC'~(m^3/m^3)))) +
    
        # Add theme
        ggplot.theme

```











