---
title: "Comparison Among different Sensor Types"
author: "James Mickley"
output:
  html_notebook:
    theme: readable
  html_document:
    keep_md: yes
    theme: readable
graphics: yes
---

```{r setup, include = F}
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ Knitr Options
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

# Set root directory to the project directory
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
?second

# Set default knitr options: 
# Suppress warnings and messages, cache chunks, 
#  set default figure size to 6x8 at 300 dpi, and save a png and pdf
knitr::opts_chunk$set(warning = F, message = F, collapse = T, cache = T,
    fig.height = 6, fig.width = 8, dpi = 300, # 6x8" @ 300dpi:1800x2400=4.3MP
    dev = c('png', 'pdf'), dev.args = list(pdf = list(onefile = F)))

```



### Overview

This analysis compares the results for the EMU, Hobo Pendant, Hobo Microstation, and iButton at the first site in each of the two transects.  


```{r 'Main_Code', include = F, cache = F}

# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ Setup - This code is run, but output is hidden
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

# Load Packages
require(cowplot) # Needed for publication-quality ggplots
require(tidyr) # Needed for data wrangling
require(lubridate) # Needed for date wrangling
require(dplyr) # Needed for data wrangling, load last to avoid conflicts

# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ Data Preparation
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

# Import datasets
data <- read.csv("Data/EMU-6-21/fentondata-all.csv")


# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ ggPlot Theme
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

ggplot.theme <- theme(
    
    # Text size for axis ticks
    axis.text.y = element_text(size = 14),
    axis.text.x = element_text(size = 14),
    
    # Text size for axis labels
    # Also move them away from the axes a bit for more space
    axis.title.x = element_text(size = 18, face = "bold", vjust = -1),
    axis.title.y = element_text(size = 18, face = "bold", vjust = 1.5),
    
    # Plot title size
    plot.title = element_text(size = 20, face = "bold"),
    
    # Margins for top, right, bottom, left
    plot.margin = grid::unit(c(1.5, 1.5, 1.5, 1.2), "lines"), 
    
    # Legend text size
    legend.text = element_text(size = 14),
    legend.text.align = 0, 
    legend.title = element_text(size = 16, face = "bold"),
    legend.key.size = grid::unit(1.4, "line"),
    legend.key = element_blank()
    )


```


```{r 'Data_Wrangling', echo = F, include = F}

# Data wrangling on the main dataset
data <- data %>%
    
    # Convert the timestamp to a date/time object
    mutate(timestamp = as.POSIXct(timestamp)) %>%
    
    # Remove the seconds from the time, leaving only hour/minute
    mutate(timestamp = timestamp - second(timestamp))

# Take a look at the structure of the main dataset
str(data)


# Convert the dataset to wide format 
data.wide <- data %>%

    # Go to pure long format: all measurements in one column
    gather(measurement, value, temperature, humidity, vwc, 
        par, light, soil, pressure) %>%
    
    # Make a column that is both source & measurement type
    unite(measure, source, measurement) %>%
    
    # Get rid of columns that aren't needed
    # Or won't work with wide format
    select(-site, -voltage, -type, -emu) %>%
    
    # Filter out the rows where the value is NA 
    filter(!is.na(value)) %>%
    
    # Go from long format to wide: each row is a site at one time
    spread(measure, value)


# Take a look at the structure of the wide dataset
str(data.wide)

```

### Temperature

Differences in temperature between EMUs and Hobos at all positions, and between EMUs and iButtons at the first position of each transect


```{r 'Temp_Diff', echo = F}


# Plot relative differences in temperature
data.wide %>%
    
    # Make columns for differences between EMUs and other sensors
    mutate(iButton_diff = iButton_temperature - EMU_temperature, 
           Hobo_diff = Hobo_temperature - EMU_temperature) %>%
    
    # Gather the differences into one column
    gather(difference, value, iButton_diff, Hobo_diff) %>%
    
    # Plot differences over time
    ggplot(aes(x = timestamp, y = value, color = difference)) +
    
        # Make a facet for each transect
        facet_wrap(~ transect, ncol = 1, scales = "free_y") + 
    
        # Add points
        geom_point(alpha = 0.5, size = 1.3) +
    
        # Add horizontal line at zero
        geom_hline(yint = 0) + 
    
        # Theme the legend
        #scale_color_manual(name = "Difference", values = c(2, 3)) + 
    
        # Add axis labels
        xlab("Date") + 
        ylab("Temperature (ºC)") + 
    
        # Add theme
        ggplot.theme

```


### Humidity

Differences in humidity between EMUs and iButtons at the first position of each transect


```{r 'Humid_Diff', echo = F}


# Plot relative differences in humidity
data.wide %>%
    
    # Only include data from the first site on a transect
    filter(order == 1) %>%
    
    mutate(difference = "iButton_diff") %>%
    
    # Plot differences over time
    ggplot(aes(x = timestamp, y = iButton_humidity - EMU_humidity, color = difference)) +
    
        # Make a facet for each transect
        facet_wrap(~ transect, ncol = 1, scales = "free_y") + 
    
        # Add points
        geom_point(alpha = 0.5, size = 1.5) +
    
        # Add horizontal line at zero
        geom_hline(yint = 0) + 
    
        # Theme the legend
        scale_color_manual(name = "Difference", values = c("blue")) + 
    
        # Add axis labels
        xlab("Date") + 
        ylab("Relative Humidity (%)") + 
    
        # Add theme
        ggplot.theme

```





### PAR

Differences in PAR between EMUs and Hobo Microstations at the first position of each transect


```{r 'PAR_Diff', echo = F}

# Plot relative differences in PAR
data.wide %>%
    
    # Only include data from the first site on a transect
    filter(order == 1) %>%
    
    mutate(difference = "Hobo_diff") %>%
    
    # Plot differences over time
    ggplot(aes(x = timestamp, y = Hobo_par - EMU_par, color = difference)) +
    
        # Make a facet for each transect
        facet_wrap(~ transect, ncol = 1, scales = "free_y") + 
    
        # Add points
        geom_point(alpha = 0.5, size = 1.5) +
    
        # Add horizontal line at zero
        geom_hline(yint = 0) + 
    
        # Theme the legend
        scale_color_manual(name = "Difference", values = c("darkgreen")) + 
    
        # Add axis labels
        xlab("Date") + 
        ylab("PAR") + 
    
        # Add theme
        ggplot.theme

```




### Volumetric Water Content

Differences in Volumetric Water Content between EMUs and Hobo Microstations at the first position of each transect


```{r 'VWC_Diff', echo = F}


# Plot relative differences in PAR
data.wide %>%
    
    # Only include data from the first site on a transect
    filter(order == 1) %>%
    
    mutate(difference = "Hobo_diff") %>%
    
    # Plot differences over time
    ggplot(aes(x = timestamp, y = Hobo_vwc - EMU_vwc, color = difference)) +
    
        # Make a facet for each transect
        facet_wrap(~ transect, ncol = 1, scales = "free_y") + 
    
        # Add points
        geom_point(alpha = 0.5, size = 1.5) +
    
        # Add horizontal line at zero
        geom_hline(yint = 0) + 
    
        # Theme the legend
        scale_color_manual(name = "Difference", values = c("black")) + 
    
        # Add axis labels
        xlab("Date") + 
        ylab("Volumetric Water Content (ml/cm^3)") + 
    
        # Add theme
        ggplot.theme

```


