---
title: "Comparison Among Sensor Types"
author: "James Mickley"
output:
  html_document:
    keep_md: yes
    theme: readable
  html_notebook:
    theme: readable
graphics: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include = F}
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ Knitr Options
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

# Set root directory to the project directory
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

# Set default knitr options: 
# Suppress warnings and messages, cache chunks, 
#  set default figure size to 6x8 at 300 dpi, and save a png and pdf
knitr::opts_chunk$set(warning = F, message = F, collapse = T, cache = T,
    fig.height = 6, fig.width = 8, dpi = 300, # 6x8" @ 300dpi:1800x2400=4.3MP
    dev = c('png', 'pdf'), dev.args = list(pdf = list(onefile = F)))

```



### Overview

This analysis compares the results for the EMU, Hobo Pendant, and iButton at the first site in each of the two transects.  


```{r 'Main_Code', include = F, cache = F}

# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ Setup - This code is run, but output is hidden
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

# Load Packages
require(tidyverse) # Needed for data wrangling
require(cowplot) # Needed for publication-quality ggplots
require(ggpmisc) # Needed for lm equations on graphs
require(lubridate) # Needed for date wrangling
require(rgeos) # Required for dawn/dusk times
require(maptools) # Required for dawn/dusk times

# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ Data Preparation
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

# Import datasets
data <- read.csv("Data/EMU-6-21/fentondata-all.csv")

# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ ggPlot Theme
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

ggplot.theme <- theme(
    
    # Text size for axis ticks
    axis.text.y = element_text(size = 14),
    axis.text.x = element_text(size = 14),
    
    # Text size for axis labels
    # Also move them away from the axes a bit for more space
    axis.title.x = element_text(size = 18, face = "bold", vjust = -1),
    axis.title.y = element_text(size = 18, face = "bold", vjust = 1.5),
    
    # Plot title size
    plot.title = element_text(size = 20, face = "bold"),
    
    # Margins for top, right, bottom, left
    plot.margin = grid::unit(c(1.5, 1.5, 1.5, 1.2), "lines"), 
    
    # Legend text size
    legend.text = element_text(size = 14),
    legend.text.align = 0, 
    legend.title = element_text(size = 16, face = "bold"),
    legend.key.size = grid::unit(1.4, "line"),
    legend.key = element_blank()
    )


```


```{r 'Data_Wrangling', echo = F, include = F}


# Data wrangling on the main dataset
data <- data %>%
    
    # Convert the timestamp to a date/time object
    mutate(timestamp = as.POSIXct(timestamp)) %>%
    
    # Remove the seconds from the time, leaving only hour/minute
    mutate(timestamp = timestamp - second(timestamp))

# Take a look at the structure of the main dataset
str(data)


# Convert the dataset to wide format 
data.wide <- data %>%

    # Go to pure long format: all measurements in one column
    gather(measurement, value, temperature, humidity, vwc, 
        pfd, light, soil, pressure) %>%
    
    # Make a column that is both source & measurement type
    unite(measure, source, measurement) %>%
    
    # Get rid of columns that aren't needed
    # Or won't work with wide format
    select(-site, -voltage, -type, -emu) %>%
    
    # Filter out the rows where the value is NA 
    filter(!is.na(value)) %>%
    
    # Go from long format to wide: each row is a site at one time
    spread(measure, value)


# Take a look at the structure of the wide dataset
str(data.wide)


# Dawn/Dusk times dataset
# See: http://rstudio-pubs-static.s3.amazonaws.com/17282_9510673f07294de7905e11c7f2b043a5.html

# Create a coordinate for the Fenton site
fenton.coord <- matrix(c(-72.238159, 41.825076), nrow = 1) %>%
    SpatialPoints(proj4string = CRS("+proj=longlat +datum=WGS84"))

# Create a data frame with dawn and dusk times with the following start/end dates for the project
darkness <- data.frame(date = seq(from = as.POSIXct("2017-05-30"), 
     to = as.POSIXct("2017-06-21"), by = "day")) %>%
     
     # Add dawn times
     mutate(dawn.time = crepuscule(fenton.coord, date, solarDep = 6, 
         direction = "dawn", POSIXct.out = TRUE)$time) %>%
  
     # Add dusk times
     mutate(dusk.time = crepuscule(fenton.coord, date, solarDep = 6, 
         direction = "dusk", POSIXct.out = TRUE)$time) %>%
    
     # Convert dawn and dusk times to hours since the start
     mutate(
         dusk = round(interval(as.POSIXct("2017-05-30 12:00"), lag(dusk.time)) /
            dhours(1)),
         dawn = round(interval(as.POSIXct("2017-05-30 12:00"), dawn.time) /
            dhours(1))) %>%
    
    # Get rid of the first day
    filter(date >= as.POSIXct("2017-05-31"))

diff.darkness <- rbind(
    darkness %>% mutate(measurement = "Temperature", ymin = -14, ymax = 8), 
    darkness %>% mutate(measurement = "Humidity", ymin = -26, ymax = 37), 
    darkness %>% mutate(measurement = "PFD", ymin = -1665, ymax = 10),
    darkness %>% mutate(measurement = "VWC", ymin = -0.18, ymax = 0)) %>%
    
    # Order the measurements
    mutate(measurement = factor(measurement, 
        levels = c("Temperature", "Humidity", "PFD", "VWC")))


```

### Temperature

Differences in temperature between EMUs relative to Hobo Pendants at all positions in the transects.

Hobos tend to read high, especially during the day.  Strongly suggests heat trapped from sun.

```{r 'Temp_Diff_Hobo', echo = F}

# Set ymin and ymax limits separately for the two transects
temp.hobo.darkness <- rbind(
    darkness %>% mutate(transect = "Meadow", ymin = -6, ymax = 10),
    darkness %>% mutate(transect = "Woods", ymin = -4, ymax = 4))

# Plot relative differences in temperature
data.wide %>%
    
    # Make columns for differences between Hobo & EMU
    mutate(Hobo_diff = Hobo_temperature - EMU_temperature) %>%
    
    # Get rid of missing data rows
    filter(!is.na(Hobo_diff)) %>%
    
    # Plot differences over time
    ggplot(aes(x = hour, y = Hobo_diff)) +
    
        # Make a facet for each transect
        facet_wrap(~ transect, ncol = 1, scales = "free_y") + 
    
        # Add rectangles for darkness
        geom_rect(data = temp.hobo.darkness, aes(xmin = dusk, xmax = dawn, 
            ymin = ymin, ymax = ymax, x = NULL, y = NULL, color = NULL), 
            fill = "grey85", show.legend = F) + 
    
        # Add points
        geom_point(alpha = 0.3) +
    
        # Add horizontal line at zero
        geom_hline(yintercept = 0, color = "red") + 
    
        # Add axis labels
        xlab("Hour") + 
        ylab("Temperature Difference (ºC)") + 
    
        # Title
        ggtitle("Hobo Temperature Difference") +
    
        # Add theme
        ggplot.theme

```

Differences in temperature between EMUs relative to iButtons at the first position of each transect.

Prior to hour 98, the iButtons had fallen on the ground, so you are seeing the difference between ground temperature and that of 10 cm above.

There is minimal difference in temperature between the iButtons and Hobos on both transects from hour 98 to 190.  After 190, the meadow BME280 started failing. Subsequent values may be inaccurate.

```{r 'Temp_Diff_iButton', echo = F}

# Set ymin and ymax limits separately for the two transects
temp.iButton.darkness <- rbind(
    darkness %>% mutate(transect = "Meadow", ymin = -8, ymax = 16),
    darkness %>% mutate(transect = "Woods", ymin = -8, ymax = 5))

# Plot relative differences in temperature
data.wide %>%
    
    # Make a colummn for differences between iButton & EMU
    mutate(iButton_diff = iButton_temperature - EMU_temperature) %>%

    # Get rid of missing data rows
    filter(!is.na(iButton_diff)) %>%
    
    # Plot differences over time
    ggplot(aes(x = hour, y = iButton_diff)) +
    
        # Make a facet for each transect
        facet_wrap(~ transect, ncol = 1, scales = "free_y") + 
    
        # Add rectangles for darkness
        geom_rect(data = temp.iButton.darkness, aes(xmin = dusk, xmax = dawn, 
            ymin = ymin, ymax = ymax, x = NULL, y = NULL, color = NULL), 
            fill = "grey85", show.legend = F) + 
    
        # Add points
        geom_point(alpha = 0.3) +
    
        # Add horizontal line at zero
        geom_hline(yintercept = 0, color = "red") + 

        # Add axis labels
        xlab("Hour") + 
        ylab("Temperature Difference (ºC)") + 
    
        # Title
        ggtitle("iButton Temperature Difference") +
    
        # Add theme
        ggplot.theme

```


### Humidity

Differences in humidity between EMUs relative to iButtons at the first position of each transect.

Prior to hour 98, the iButtons had fallen on the ground, so this is the difference between ground temperature and that of 10 cm above.

It looks like humidity is lower during the day with the iButton on the meadow transect, and generally higher at all times in the woods.


```{r 'Humid_Diff', echo = F}

# Set ymin and ymax limits separately for the two transects
humid.darkness <- rbind(
    darkness %>% mutate(transect = "Meadow", ymin = -36, ymax = 55),
    darkness %>% mutate(transect = "Woods", ymin = -12, ymax = 72))


# Plot relative differences in humidity
data.wide %>%
    
    # Only include data from the first site on a transect
    #filter(order == 1, hour > 98) %>%
    
    # Make a colummn for differences between iButton & EMU
    mutate(iButton_diff = iButton_humidity - EMU_humidity) %>%

    # Get rid of missing data rows
    filter(!is.na(iButton_diff)) %>%
    
    # Plot differences over time
    ggplot(aes(x = hour, y = iButton_diff)) +
    
        # Make a facet for each transect
        facet_wrap(~ transect, ncol = 1, scales = "free_y") + 

        # Add rectangles for darkness
        geom_rect(data = humid.darkness, aes(xmin = dusk, xmax = dawn, 
            ymin = ymin, ymax = ymax, x = NULL, y = NULL, color = NULL), 
            fill = "grey85", show.legend = F) + 
    
        # Add points
        geom_point(alpha = 0.3) +
    
        # Add horizontal line at zero
        geom_hline(yintercept = 0, color = "red") + 
    
        # Theme the legend
        scale_color_manual(name = "Difference", values = c("blue")) + 
    
        # Add axis labels
        xlab("Date") + 
        ylab("Relative Humidity (%)") + 
    
        # Title
        ggtitle("iButton Humidity Difference") +
    
        # Add theme
        ggplot.theme

```




## Comparisons between EMU & iButton/Hobo Pendant for Woods transect

The Hobo Pendant reads slightly high on temperature (by about 2ºC).  Presumably it's getting heated inside of its enclosed case.


```{r 'Temp_Comp_Hobo', echo = F, fig.width=10}
    
# Plot relative differences in temperature
data.wide %>%
    
    # Get rid of the first position in the meadow transect
    filter(!is.na(EMU_temperature), !(transect == "Meadow" & order == 1)) %>%
    
    # Plot differences over time
    ggplot(aes(y = EMU_temperature, x = Hobo_temperature)) +
    
        # Make a facet for each transect
        facet_wrap(~ transect, ncol = 2) + 
    
        # Add points
        geom_point(color = "blue", alpha = 0.6) +
    
        # Add horizontal line at zero
        geom_abline(intercept = 0, slope = 1) + 
    
        # Add linear model
        geom_smooth(method = "lm", color = "red", se = F) + 
    
        # Add the equation
        stat_poly_eq(aes(label = ..eq.label..),
               eq.x.rhs = "italic(x)", coef.digits = 3, label.x = 15, 
               label.y = 40, parse = T, size = 7,
               formula = y ~ x) +
        
        # Add the R^2
        stat_poly_eq(aes(label = ..rr.label..),
               rr.digits = 3, label.x = 15, label.y = 36,
               parse = T, size = 7, formula = y ~ x) +
    
        # Add horizontal line at zero
        geom_abline(intercept = 0, slope = 1) + 
    
        # Add axis labels
        xlab("Hobo Temperature (ºC)") + 
        ylab("EMU Temperature (ºC)") + 
    
        # Add theme
        ggplot.theme



# Combine transects into one graph
hobo.temp <- data.wide %>%
    
    # Get rid of the first position in the meadow transect
    filter(!is.na(EMU_temperature), !(transect == "Meadow" & order == 1)) %>%
    
    # Plot differences over time
    ggplot(aes(y = EMU_temperature, x = Hobo_temperature)) +
    
        # Add points
        geom_point(color = "blue", alpha = 0.6) +
    
        # Add horizontal line at zero
        geom_abline(intercept = 0, slope = 1) + 
    
        # Add linear model
        geom_smooth(method = "lm", color = "red", se = F) + 
    
        # Add the equation
        stat_poly_eq(aes(label = ..eq.label..),
               eq.x.rhs = "italic(x)", coef.digits = 3, label.x = 15, 
               label.y = 46, parse = T, size = 7,
               formula = y ~ x) +
        
        # Add the R^2
        stat_poly_eq(aes(label = ..rr.label..),
               rr.digits = 3, label.x = 15, label.y = 40,
               parse = T, size = 7, formula = y ~ x) +
    
        # Add horizontal line at zero
        geom_abline(intercept = 0, slope = 1) + 
    
        # Add axis labels
        xlab("Hobo Temperature (ºC)") + 
        ylab("EMU Temperature (ºC)") + 
    
        # Add theme
        ggplot.theme


```

The iButtons read slightly lower than the EMU when it's cold, and slightly higher when hot.  Presumably, this is in response to the metal enclosure.

```{r 'Temp_Comp_iButton', echo = F, fig.width=10}

# Plot relative differences in temperature
data.wide %>%
    
    # Only show the woods transect at the first site
    filter(order == 1, hour > 98, 
        !is.na(EMU_temperature), !is.na(iButton_temperature)) %>%
    
    # Plot differences over time
    ggplot(aes(y = EMU_temperature, x = iButton_temperature)) +

        # Make a facet for each transect
        facet_wrap(~ transect, ncol = 2) + 
    
        # Add points
        geom_point(color = "blue", alpha = 0.6) +
    
        # Add horizontal line at zero
        geom_abline(intercept = 0, slope = 1) + 
    
        # Add linear model
        geom_smooth(method = "lm", color = "red", se = F) + 
    
        # Add the equation
        stat_poly_eq(aes(label = ..eq.label..),
               eq.x.rhs = "italic(x)", coef.digits = 4, label.x = 15, 
               label.y = 40, parse = T, size = 7,
               formula = y ~ x) +
        
        # Add the R^2
        stat_poly_eq(aes(label = ..rr.label..),
               rr.digits = 3, label.x = 15, label.y = 36,
               parse = T, size = 7, formula = y ~ x) +
    
        # Add horizontal line at zero
        geom_abline(intercept = 0, slope = 1) + 
    
        # Theme the legend
        #scale_color_manual(name = "Difference", values = c(2, 3)) + 
    
        # Add axis labels
        xlab("iButton Temperature (ºC)") + 
        ylab("EMU Temperature (ºC)") + 
    
        # Add theme
        ggplot.theme



# Plot relative differences in temperature
ibutton.temp <- data.wide %>%
    
    # Only show the woods transect at the first position after hour 98
    filter(!is.na(EMU_temperature), !is.na(iButton_temperature), 
           order == 1, hour > 98, transect != "Meadow") %>%
    
    # Plot differences over time
    ggplot(aes(y = EMU_temperature, x = iButton_temperature)) +
    
        # Add points
        geom_point(color = "blue", alpha = 0.6) +
    
        # Add horizontal line at zero
        geom_abline(intercept = 0, slope = 1) + 
    
        # Add linear model
        geom_smooth(method = "lm", color = "red", se = F) + 
    
        # Add the equation
        stat_poly_eq(aes(label = ..eq.label..),
               eq.x.rhs = "italic(x)", coef.digits = 4, label.x = 16.5, 
               label.y = 32, parse = T, size = 7,
               formula = y ~ x) +
        
        # Add the R^2
        stat_poly_eq(aes(label = ..rr.label..),
               rr.digits = 3, label.x = 16.5, label.y = 29,
               parse = T, size = 7, formula = y ~ x) +
    
        # Add horizontal line at zero
        geom_abline(intercept = 0, slope = 1) + 
    
        # Theme the legend
        #scale_color_manual(name = "Difference", values = c(2, 3)) + 
    
        # Add axis labels
        xlab("iButton Temperature (ºC)") + 
        ylab("EMU Temperature (ºC)") + 
    
        # Add theme
        ggplot.theme


```

Comparing iButtons to Hobos, we see that hobos read too high for anything over about 10C.  Again evidence of self-heating.

```{r 'Temp_Comp_iButton_Hobo', echo = F, fig.width=10}

# Plot relative differences in temperature between Hobos and iButtons
hobo.ibutton.temp <- data.wide %>%
    
    # Only show the first position after hour 98 (when iButtons fixed)
    filter(!is.na(Hobo_temperature), !is.na(iButton_temperature), 
           order == 1, hour > 98) %>%
    
    # Plot differences over time
    ggplot(aes(y = Hobo_temperature, x = iButton_temperature)) +
    
        # Add points
        geom_point(color = "blue", alpha = 0.6) +
    
        # Add horizontal line at zero
        geom_abline(intercept = 0, slope = 1) + 
    
        # Add linear model
        geom_smooth(method = "lm", color = "red", se = F) + 
    
        # Add the equation
        stat_poly_eq(aes(label = ..eq.label..),
               eq.x.rhs = "italic(x)", coef.digits = 4, label.x = 13, 
               label.y = 44, parse = T, size = 7,
               formula = y ~ x) +
        
        # Add the R^2
        stat_poly_eq(aes(label = ..rr.label..),
               rr.digits = 3, label.x = 13, label.y = 38,
               parse = T, size = 7, formula = y ~ x) +
    
        # Add horizontal line at zero
        geom_abline(intercept = 0, slope = 1) + 
    
        # Theme the legend
        #scale_color_manual(name = "Difference", values = c(2, 3)) + 
    
        # Add axis labels
        xlab("iButton Temperature (ºC)") + 
        ylab("Hobo Temperature (ºC)") + 
    
        # Add theme
        ggplot.theme

# Display
hobo.ibutton.temp

```




### Humidity

The iButtons have slightly higher humidity (3-5% or so) in the Woods.  They also can return values higher than 100%.  

In the meadow, the BME280 failed, and I think all the humidity values are bogus.

```{r 'Humid_Comp', echo = F, fig.width=10}


# Plot relative differences in humidity
data.wide %>%

    # Only show the woods transect at the first position after hour 98
    filter(!is.na(EMU_humidity), !is.na(iButton_humidity), 
           order == 1, hour > 98) %>%
    
    # Plot differences over time
    ggplot(aes(x = iButton_humidity, y = EMU_humidity)) +

        # Make a facet for each transect
        facet_wrap(~ transect, ncol = 2) + 

        # Add points
        geom_point(color = "blue", alpha = 0.6) +
    
        # Add horizontal line at zero
        geom_abline(intercept = 0, slope = 1) + 

        # Add linear model
        geom_smooth(method = "lm", color = "red", se = F) + 
    
        # Add the equation
        stat_poly_eq(aes(label = ..eq.label..),
               eq.x.rhs = "italic(x)", coef.digits = 4, label.x = 50, 
               label.y = 100, parse = T, size = 7,
               formula = y ~ x) +
        
        # Add the R^2
        stat_poly_eq(aes(label = ..rr.label..),
               rr.digits = 3, label.x = 50, label.y = 95,
               parse = T, size = 7, formula = y ~ x) +
    
        # Theme the legend
        #scale_color_manual(name = "Difference", values = c(2, 3)) + 
    
        # Add axis labels
        xlab("iButton Humidity (%)") + 
        ylab("EMU Humidity (%)") + 
    
        # Add theme
        ggplot.theme





# Plot relative differences in humidity
ibutton.humid <- data.wide %>%

    # Only show the woods transect at the first position after hour 98
    filter(!is.na(EMU_humidity), !is.na(iButton_humidity), 
           order == 1, hour > 98, transect != "Meadow") %>%
    
    # Plot differences over time
    ggplot(aes(x = iButton_humidity, y = EMU_humidity)) +

        # Add points
        geom_point(color = "blue", alpha = 0.6) +
    
        # Add horizontal line at zero
        geom_abline(intercept = 0, slope = 1) + 

        # Add linear model
        geom_smooth(method = "lm", color = "red", se = F) + 
    
        # Add the equation
        stat_poly_eq(aes(label = ..eq.label..),
               eq.x.rhs = "italic(x)", coef.digits = 3, label.x = 56, 
               label.y = 100, parse = T, size = 7,
               formula = y ~ x) +
        
        # Add the R^2
        stat_poly_eq(aes(label = ..rr.label..),
               rr.digits = 3, label.x = 56, label.y = 90,
               parse = T, size = 7, formula = y ~ x) +
    
        # Theme the legend
        #scale_color_manual(name = "Difference", values = c(2, 3)) + 
    
        # Add axis labels
        xlab("iButton Humidity (%)") + 
        ylab("EMU Humidity (%)") + 
    
        # Add theme
        ggplot.theme

```

```{r 'Temp_Humid_Comp', echo = F, fig.height=10, fig.width=10}

ggdraw() +
    
    # Add the four temp and humidity comparison graphs
    draw_plot(hobo.temp, 0, 0.49, 0.5, 0.49) +
    draw_plot(ibutton.temp, 0.5, 0.49, 0.5, 0.49) +
    draw_plot(hobo.ibutton.temp, 0.5, 0, 0.5, 0.49) +
    draw_plot(ibutton.humid, 0, 0, 0.5, 0.49) + 
    
    # Add subplot labels
    draw_plot_label(c("A", "B", "C", "D"), 
        x = c(0.01, 0.51, 0.01, 0.51), 
        y = c(1, 1, 0.51, 0.51), size = 24)


```

### Session Information

```{r 'Session_Info', echo = F, comment = ""}

# Add session information to help with reproduceability
sessionInfo()


```
