---
title: "Calibration"
author: "James Mickley"
output:
  html_document:
    keep_md: yes
    theme: readable
    toc: true
  html_notebook:
    theme: readable
graphics: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include = F}
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ Knitr Options
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

# Set root directory to the project directory
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())


# Set default knitr options: 
# Suppress warnings and messages, cache chunks, 
#  set default figure size to 6x8 at 300 dpi, and save a png and pdf
knitr::opts_chunk$set(warning = F, message = F, collapse = T, cache = T,
    fig.height = 6, fig.width = 8, dpi = 300, # 6x8" @ 300dpi:1800x2400=4.3MP
    dev = c('png', 'pdf'), dev.args = list(pdf = list(onefile = F)))

```



### Overview

This analysis calibrates the light and soil moisture sensors.

The light sensor is calibrated for PFD against a LiCor and compared to a Hobo, while the soil probes are calibrated against measured volumetric water content and compared to a Hobo.

```{r 'Main_Code', include = F, cache = F}

# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ Setup - This code is run, but output is hidden
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

# Load Packages
require(tidyverse) # Needed for ggplot2 and data wrangling
require(cowplot) # Needed for publication-quality ggplots
require(ggpmisc) # Needed for lm equations on graphs
require(bbmle) # Needed for AICctab

# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ Data Preparation
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

# Import datasets
light <- read.csv("Data/licor-calibration/licor-calibration.csv")
soil <- read.csv("Data/Fenton-soil/Fenton_soil_calibration.csv")


# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ ggPlot Theme
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

ggplot.theme <- theme(
    
    # Text size for axis ticks
    axis.text.y = element_text(size = 14),
    axis.text.x = element_text(size = 14),
    
    # Text size for axis labels
    # Also move them away from the axes a bit for more space
    axis.title.x = element_text(size = 18, face = "bold", vjust = -1),
    axis.title.y = element_text(size = 18, face = "bold", vjust = 1.5),
    
    # Plot title size
    plot.title = element_text(size = 20, face = "bold"),
    
    # Margins for top, right, bottom, left
    plot.margin = grid::unit(c(1.5, 1.5, 1.5, 1.2), "lines"), 
    
    # Legend text size
    legend.text = element_text(size = 14),
    legend.text.align = 0, 
    legend.title = element_text(size = 16, face = "bold"),
    legend.key.size = grid::unit(1.4, "line"),
    legend.key = element_blank()
    )


```

### Light Calibration Results


This shows the calibration of the [BH1750FVI light sensor](http://www.mouser.com/ds/2/348/bh1750fvi-e-186247.pdf) against a [LiCor LI-190R Quantum Sensor](https://www.licor.com/env/products/light/quantum.html) connected to a [LiCor LI-250 Light Meter](https://www.licor.com/env/products/light/light_meter.html)

Both were mounted at the same angle and height on a board, and taking readings every second.

#### Light Linear Model

```{r 'Light_linear_model', echo = F, comment = ""}

# Run a linear model
light.lm <- lm(umol ~ lux, data = light)

# Show summary
summary(light.lm)

```

#### Light Quadratic Model

```{r 'Light_quadratic_model', echo = F, comment = ""}

# Run a quadratic model
light.quad <- light %>%

    # Add a squared column
    mutate(lux.squared = lux ^ 2) %>%
    
    # model
    lm(umol ~ lux + lux.squared, data = .)

# Show summary
summary(light.quad)

# Rename the intercept coefficient
names(light.quad$coefficients)[1] <- "intercept"

# Write the calibration coefficients to a CSV
write.csv(data.frame(as.list(light.quad$coefficients)),
    "Analyses/Calibration/light-calibration.csv", row.names = F)

```


```{r 'Predict_Light', echo = F, include = F}

# Update the dataset with predicted values from the linear models
light <- light %>%
    
    # Add the linear calibration
    mutate(umol.linear = light.lm$coefficients[1] + 
        light.lm$coefficients[2] * lux) %>%
    
    # Add the quadratic calibration
    mutate(umol.quad = light.quad$coefficients[1] + 
        light.quad$coefficients[2] * lux + 
        light.quad$coefficients[3] * lux ^ 2)

```



#### Light Calibration Graphs

Comparison between the linear and quadratic models

```{r 'Light_model_comparison', echo = F}

light %>%
    
    # Make a column with both the linear and quad predicted values
    gather(calibration, value, umol.linear, umol.quad) %>%
    
    # Construct the graph
    ggplot(aes(x = umol, y = value, color = calibration)) +
    
        # Add a line with a slope of 1 and intercept of 0
        geom_abline(intercept = 0, slope = 1) + 

        # Add points
        geom_point() +
    
        # 
        scale_color_manual(name = "Model", labels = c("Linear", "Quadratic"), 
            values = c("red", "blue")) +
    
        # Add axes
        xlab(bquote(bold('LiCor PFD (µmol '~m^-2 ~s^-1~')'))) +
        ylab(bquote(bold('Calibrated PFD (µmol '~m^-2 ~s^-1~')'))) +
    
        # Add theme
        ggplot.theme + 
    
        # Move the legend
        theme(legend.position = c(0.2, 0.85))

```


The best calibration is the quadratic model, presented here alone


```{r 'Light_quad_calibration', echo = F}

# Construct the graph
light %>% ggplot(aes(x = lux, y = umol)) +
    
        # Add points
        geom_point() +
    
        # Add linear model
        geom_smooth(method = "lm", color = "red", se = F, 
            formula = y ~ poly(x, 2, raw = T)) + 
    
        # Add the equation
        stat_poly_eq(aes(label = ..eq.label..),
               eq.x.rhs = "italic(x)", coef.digits = 2, label.x = 50000, 
               label.y = 2000, parse = T, size = 7,
               formula = y ~ poly(x, 2, raw = T)) +
        
        # Add the R^2
        stat_poly_eq(aes(label = ..rr.label..),
               coef.digits = 2, label.x = 50000, label.y = 1800,
               parse = T, size = 7, formula = y ~ poly(x, 2, raw = T)) +
    
        # Add axes
        ylab(bquote(bold('LiCor PFD (µmol '~m^-2 ~s^-1~')'))) +
        xlab("Sensor Value (Lux)") +
    
        # Add theme
        ggplot.theme

```

[Root mean square](https://en.wikipedia.org/wiki/Root_mean_square#Error) error between predicted values from the quadratic light model and PFD values from the LiCor.  This gives a measure of accuracy.

```{r 'RMSE_Light', echo = F, comment = ""}

# Calculate the root mean square of the residuals
rms.light <- light %>%
    
    # Calculate the difference between the LiCor and the prediction
    mutate(umol.diff = umol - umol.quad) %>%
    
    # Square the differences
    mutate(diff.squared = umol.diff ^ 2) %>%
    
    # Calculate the mean square and root mean square
    summarize(mean.square = mean(diff.squared), 
        root.mean.square = sqrt(mean.square))

# Display
rms.light

```

Graph showing the predicted values of PFD from the quadratic model graphed against the LiCor

```{r 'Light_predict', echo = F}

#TODO: Add equation to graph

# Construct the graph
light %>% ggplot(aes(x = umol, y = umol.quad)) +
    
        # Add a line with a slope of 1 and intercept of 0
        geom_abline(intercept = 0, slope = 1) + 
    
        # Add points
        geom_point() +
    
        # Add axes
        xlab(bquote(bold('LiCor PFD (µmol '~m^-2 ~s^-1~')'))) +
        ylab(bquote(bold('Predicted Sensor PFD (µmol '~m^-2 ~s^-1~')'))) +
    
        # Add theme
        ggplot.theme

```


### Soil Calibration Results

Calibration of generic silver and gold soil moisture sensor probes with a datalogging scale, measuring water content of the soil every five minutes. The soil used for calibration was a rich loam from the Fenton Meadow. The data columns soil1 and soil2 are gold (ENIG) probes, while soil3 is a silver (HASL) probe.


```{r 'Data_Wrangling_Soil', echo = F, comment = ""}

# From the metadata
fenton.dryweight = 992.7
fenton.wetweight = 1638.7
fenton.density = 0.90

# Data wrangling
cal <- soil %>%

    # Remove the metadata column
    select(-metadata) %>%
    
    # Get rid of any missing data rows
    filter(!is.na(weight)) %>%
    
    # Add columns for measured gravimetric and volumetric water content
    mutate(gwc = (weight - fenton.dryweight) / fenton.dryweight, 
        vwc = gwc * fenton.density) %>%
    
    # Convert the date to date format
    mutate(date = as.POSIXct(date, format = "%m/%d/%Y %H:%M"))


```

### Soil Gold Models

Models using the data from the gold-plated soil moisture probe

#### Soil Gold Linear Model

```{r 'Soil_linear_gold_model', echo = F, comment = ""}


# Calibrate the gold probes using a simple linear model
gold.lm <- cal %>%
    
    # Filter for after 7/22, when probes hit a minimum
    filter(date >= as.POSIXct("2017-07-22 00:00")) %>%
    
    # Run the linear model
    lm(vwc ~ gold, data = .)

# View the model
summary(gold.lm)

```

#### Soil Gold Quadratic Model

The quadratic model is considerably better

```{r 'Soil_quadratic_gold_model', echo = F, comment = ""}


# Calibrate the gold probes using a quadratic model
gold.quad <- cal %>%
  
    # Filter for after 7/22, when probes hit a minimum
    filter(date >= as.POSIXct("2017-07-22 00:00")) %>%
    
    # Make a squared column
    mutate(gold.squared = gold ^ 2) %>%

    # Run the quadratic model
    lm(vwc ~ gold + gold.squared, data = .)
 
# View the model
summary(gold.quad)

# Check AICsh
AICctab(gold.lm, gold.quad)

# Rename the intercept coefficient
names(gold.quad$coefficients) <- c("intercept", "soil", "soil.squared")

# Write the calibration coefficients to a CSV
write.csv(data.frame(as.list(gold.quad$coefficients)),
    "Analyses/Calibration/gold-soil-calibration.csv", row.names = F)

```


#### Soil Gold Quadratic Temperature Model

The quadratic temperature-compensated model is better by AIC, but it doesn't add much, and doesn't eliminate the residual zig-zag pattern (see below).

```{r 'Soil_quadratic_gold_temp_model', echo = F, comment = ""}


# Calibrate the gold probes using a quadratic model
gold.quad.temp <- cal %>%
  
    # Filter for after 7/22, when probes hit a minimum
    filter(date >= as.POSIXct("2017-07-22 00:00")) %>%
    
    # Make a squared column
    mutate(gold.squared = gold ^ 2) %>%

    # Run the quadratic model
    lm(vwc ~ gold + gold.squared + temp, data = .)
 
# View the model
summary(gold.quad.temp)

# Check AICsh
AICctab(gold.lm, gold.quad, gold.quad.temp)


```


### Soil Gold Models

Models using the data from the silver (solder plated) soil moisture probe

#### Soil Silver Linear Model

```{r 'Soil_linear_silver_model', echo = F, comment = ""}


# Calibrate the silver probe using a simple linear model
silver.lm <- cal %>%
    
    # Filter for after 7/22, when probes hit a minimum
    filter(date >= as.POSIXct("2017-07-22 00:00")) %>%
    
    # Run the linear model
    lm(vwc ~ silver, data = .)

# View the model
summary(silver.lm)

```

#### Soil Silver Quadratic Model

```{r 'Soil_quadratic_silver_model', echo = F, comment = ""}


# Calibrate the silver probe using a quadratic model
silver.quad <- cal %>%
    
    # Filter for after 7/22, when probes hit a minimum
    filter(date >= as.POSIXct("2017-07-22 00:00")) %>%
    
    # Make a squared column
    mutate(silver.squared = silver ^ 2) %>%
    
    # Run the linear model
    lm(vwc ~ silver + silver.squared, data = .)

# View the model
summary(silver.quad)

# Check AICc
AICctab(silver.lm, silver.quad)

# Rename the intercept coefficient
names(silver.quad$coefficients) <- c("intercept", "soil", "soil.squared")

# Write the calibration coefficients to a CSV
write.csv(data.frame(as.list(silver.quad$coefficients)),
    "Analyses/Calibration/silver-soil-calibration.csv", row.names = F)

```

#### Soil Silver Quadratic Temperature Model

```{r 'Soil_quadratic_silver_temp_model', echo = F, comment = ""}


# Calibrate the silver probe using a quadratic model
silver.quad.temp <- cal %>%
    
    # Filter for after 7/22, when probes hit a minimum
    filter(date >= as.POSIXct("2017-07-22 00:00")) %>%
    
    # Make a squared column
    mutate(silver.squared = silver ^ 2) %>%
    
    # Run the linear model
    lm(vwc ~ silver + silver.squared + temp, data = .)

# View the model
summary(silver.quad.temp)

# Check AICc
AICctab(silver.lm, silver.quad.temp)

```



```{r 'Predict_Soil', echo = F, include = F}

# Update the dataset with predicted values from the linear models
cal <- cal %>%
    
    # Add the quadratic calibration for the gold probe
    mutate(gold_vwc = gold.quad$coefficients[1] + 
        gold.quad$coefficients[2] * gold + 
        gold.quad$coefficients[3] * gold ^ 2) %>%
    
    # Add the temperature compensated quadratic calibration for the gold probe
    mutate(gold_vwc_temp = gold.quad.temp$coefficients[1] + 
        gold.quad.temp$coefficients[2] * gold + 
        gold.quad.temp$coefficients[3] * gold ^ 2 + 
        gold.quad.temp$coefficients[4] * temp) %>%
    
    # Add the quadratic calibration for the silver probe
    mutate(silver_vwc = silver.quad$coefficients[1] + 
        silver.quad$coefficients[2] * silver + 
        silver.quad$coefficients[3] * silver ^ 2) %>%

    # Add the temperature compensated quadratic calibration for the silver probe
    mutate(silver_vwc_temp = silver.quad.temp$coefficients[1] + 
        silver.quad.temp$coefficients[2] * silver + 
        silver.quad.temp$coefficients[3] * silver ^ 2 + 
        silver.quad.temp$coefficients[4] * temp)


```

#### Soil Calibration Graphs

Showing soil moisture probe values over time. There is some diurnal variation that is partially temperature-driven.

Adding temperature to the quadratic fit does improve things slightly, but the temperature range is only ~21-30 C, so it doesn't encompass the full temperature range outside

```{r 'Soil_probes_over_time', echo = F}

cal %>%
    
    # Filter for after 7/22, when probes hit a minimum
    filter(date >= as.POSIXct("2017-07-22 00:00")) %>%
    
    gather(probe, value, gold, silver) %>%

    ggplot(aes(x = date, y = value, color = probe, group = probe)) +

        # Set up the colors and legend
        scale_color_manual(values = c("gold", "gray60"), name = "Probe", 
            labels = c("Gold", "Silver")) +
    
        geom_line(size = 1.2) +
    
        xlab("Time") +
        ylab("Soil Probe Value") + 
    
        # Add theme
        ggplot.theme + 
    
        # Change legend position
        theme(legend.position = c(0.14, 0.85))
    

```



[Root mean square](https://en.wikipedia.org/wiki/Root_mean_square#Error) error between predicted values from the gold probe quadratic model and measured VWC values from the scale.  This gives a measure of accuracy.

```{r 'RMSE_Soil', echo = F, comment = ""}

# Calculate the root mean square of the residuals
rms.soil <- cal %>%
    
    # Calculate the difference between the actual VWC and the prediction
    mutate(vwc.diff = vwc - gold_vwc) %>%
    
    # Square the differences
    mutate(diff.squared = vwc.diff ^ 2) %>%
    
    # Calculate the mean square and root mean square
    summarize(mean.square = mean(diff.squared), 
        root.mean.square = sqrt(mean.square))

# Display
rms.soil

```


A graph of the calibration and formula using the quadratic model for the gold probe

```{r 'Soil_quad_calibration', echo = F}

cal %>%
    
    # Filter for after 7/22, when probes hit a minimum
    filter(date >= as.POSIXct("2017-07-22 00:00")) %>%

    ggplot(aes(x = gold, y = vwc)) +
    
        # Add points
        geom_point() + 
    
        # Add linear model
        geom_smooth(method = "lm", color = "red", se = F, 
            formula = y ~ poly(x, 2, raw = T)) + 
    
        # Add the equation
        stat_poly_eq(aes(label = ..eq.label..),
               eq.x.rhs = "italic(x)", coef.digits = 2, label.x = 16600, 
               label.y = 0.30, parse = T, size = 7,
               formula = y ~ poly(x, 2, raw = T)) +
        
        # Add the R^2
        stat_poly_eq(aes(label = ..rr.label..),
               coef.digits = 2, label.x = 16600, label.y = 0.28,
               parse = T, size = 7, formula = y ~ poly(x, 2, raw = T)) +
    
        # Add axis labels
        xlab("Soil Probe Value") + 
        ylab(bquote(bold('Actual VWC'~(m^3/m^3)))) +
    
        # Add theme
        ggplot.theme
    

```

Graph showing the predicted values of VWC from the quadratic model graphed against the actual VWC. Notably, the zig-zag pattern improves slightly in the temperature-compensated model, but it is far from eliminated.

```{r 'Soil_predict_comparison', echo = F}

# Plot all the EMU probes vs actual VWC
cal %>%
    
    # Filter for after 7/22, when probes hit a minimum
    filter(date >= as.POSIXct("2017-07-22 00:00")) %>%
    
    # Gather all the probes together
    gather(probe, value, gold_vwc, gold_vwc_temp, silver_vwc) %>%
    
    arrange(desc(probe)) %>%

    # Make the plot
    ggplot(aes(x = vwc, y = value, color = probe)) +
    
        # Add a line with a slope of 1 and intercept of 0
        geom_abline(intercept = 0, slope = 1) + 
    
        # Add points
        geom_point() +
    
        # Set up the colors and legend
        scale_color_manual(values = c("gold", "darkgoldenrod", "gray60"), 
            name = "Probe", labels = c("Gold", "Gold Temp", "Silver")) +
    
        # Add axis labels
        xlab(bquote(bold('Actual VWC'~(m^3/m^3)))) +
        ylab(bquote(bold('Predicted Probe VWC'~(m^3/m^3)))) + 
    
    
        # Add theme
        ggplot.theme + 
    
        # Change legend position
        theme(legend.position = c(0.14, 0.85))
 

```


### Session Information

```{r 'Session_Info', echo = F, comment = ""}

# Add session information to help with reproduceability
sessionInfo()


```
