---
title: "LiCor Calibration"
author: "James Mickley"
output:
  html_document:
    keep_md: yes
    theme: readable
    toc: true
  html_notebook:
    theme: readable
graphics: yes
---

```{r setup, include = F}
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ Knitr Options
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

# Set root directory to the project directory
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())


# Set default knitr options: 
# Suppress warnings and messages, cache chunks, 
#  set default figure size to 6x8 at 300 dpi, and save a png and pdf
knitr::opts_chunk$set(warning = F, message = F, collapse = T, cache = T,
    fig.height = 6, fig.width = 8, dpi = 300, # 6x8" @ 300dpi:1800x2400=4.3MP
    dev = c('png', 'pdf'), dev.args = list(pdf = list(onefile = F)))

```



### Overview

This analysis calibrates the light and soil moisture sensors.

The light sensor is calibrated for PAR against a LiCor and compared to a Hobo, while the soil probes are calibrated against measured volumetric water content and compared to a Hobo.

```{r 'Main_Code', include = F, cache = F}

# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ Setup - This code is run, but output is hidden
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

# Load Packages
require(cowplot) # Needed for publication-quality ggplots
require(ggpmisc) # Needed for lm equations on graphs
require(bbmle) # Needed for AICctab
require(tidyr) # Needed for data wrangling
require(dplyr) # Needed for data wrangling, load last to avoid conflicts

# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ Data Preparation
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

# Import datasets
light <- read.csv("Data/licor-calibration/licor-calibration.csv")
soil <- read.csv("Data/Fenton-soil/Fenton_soil_calibration.csv")


# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ ggPlot Theme
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

ggplot.theme <- theme(
    
    # Text size for axis ticks
    axis.text.y = element_text(size = 14),
    axis.text.x = element_text(size = 14),
    
    # Text size for axis labels
    # Also move them away from the axes a bit for more space
    axis.title.x = element_text(size = 18, face = "bold", vjust = -1),
    axis.title.y = element_text(size = 18, face = "bold", vjust = 1.5),
    
    # Plot title size
    plot.title = element_text(size = 20, face = "bold"),
    
    # Margins for top, right, bottom, left
    plot.margin = grid::unit(c(1.5, 1.5, 1.5, 1.2), "lines"), 
    
    # Legend text size
    legend.text = element_text(size = 14),
    legend.text.align = 0, 
    legend.title = element_text(size = 16, face = "bold"),
    legend.key.size = grid::unit(1.4, "line"),
    legend.key = element_blank()
    )


```

### Light Calibration Results


This shows the calibration of the [BH1750FVI light sensor](http://www.mouser.com/ds/2/348/bh1750fvi-e-186247.pdf) against a [LiCor LI-190R Quantum Sensor](https://www.licor.com/env/products/light/quantum.html) connected to a [LiCor LI-250 Light Meter](https://www.licor.com/env/products/light/light_meter.html)

Both were mounted at the same angle and height on a board, and taking readings every second.

#### Light Linear Model

```{r 'Light_linear_model', echo = F, comment = ""}

# Run a linear model
light.lm <- lm(umol ~ lux, data = light)

# Show summary
summary(light.lm)

```

#### Light Quadratic Model

```{r 'Light_quadratic_model', echo = F, comment = ""}

# Run a quadratic model
light.quad <- light %>%

    # Add a squared column
    mutate(lux.squared = lux ^ 2) %>%
    
    # model
    lm(umol ~ lux + lux.squared, data = .)

# Show summary
summary(light.quad)

# Rename the intercept coefficient
names(light.quad$coefficients)[1] <- "intercept"

# Write the calibration coefficients to a CSV
write.csv(data.frame(as.list(light.quad$coefficients)), "light-calibration.csv", 
    row.names = F)

```


```{r 'Predict_Light', echo = F, include = F}

# Update the dataset with predicted values from the linear models
light <- light %>%
    
    # Add the linear calibration
    mutate(umol.linear = light.lm$coefficients[1] + 
        light.lm$coefficients[2] * lux) %>%
    
    # Add the quadratic calibration
    mutate(umol.quad = light.quad$coefficients[1] + 
        light.quad$coefficients[2] * lux + 
        light.quad$coefficients[3] * lux ^ 2)


```


```{r 'Data_Wrangling_Light', echo = F, include = F}

# Wrangle the hobo comparison data
light.hobo.comp <- soil %>%

    # Select only lux and hobo_par (renaming to umol)
    select(lux, umol = hobo_par) %>%
    
    # Remove rows with missing data
    filter(!is.na(umol), !is.na(lux)) %>%
    
    # Add the quadratic calibration
    mutate(umol.quad = light.quad$coefficients[1] + 
        light.quad$coefficients[2] * lux + 
        light.quad$coefficients[3] * lux ^ 2) 
    
```



#### Light Calibration Graphs

Comparison between the linear and quadratic models

```{r 'Light_model_comparison', echo = F}

light %>%
    
    # Make a column with both the linear and quad predicted values
    gather(calibration, value, umol.linear, umol.quad) %>%
    
    # Construct the graph
    ggplot(aes(x = umol, y = value, color = calibration)) +
    
        # Add a line with a slope of 1 and intercept of 0
        geom_abline(intercept = 0, slope = 1) + 

        # Add points
        geom_point() +
    
        # 
        scale_color_manual(name = "Model", labels = c("Linear", "Quadratic"), 
            values = c("red", "blue")) +
    
        # Add axes
        xlab(bquote(bold('LiCor PAR (µmol '~m^-2 ~s^-1~')'))) +
        ylab(bquote(bold('Calibrated PAR (µmol '~m^-2 ~s^-1~')'))) +
    
        # Add theme
        ggplot.theme + 
    
        # Move the legend
        theme(legend.position = c(0.2, 0.85))

```


The best calibration is the quadratic model, presented here alone


```{r 'Light_quad_calibration', echo = F}

# Construct the graph
light %>% ggplot(aes(x = lux, y = umol)) +
    
        # Add points
        geom_point() +
    
        # Add linear model
        geom_smooth(method = "lm", color = "red", se = F, 
            formula = y ~ poly(x, 2, raw = T)) + 
    
        # Add the equation
        stat_poly_eq(aes(label = ..eq.label..),
               eq.x.rhs = "italic(x)", coef.digits = 2, label.x = 50000, 
               label.y = 2000, parse = T, size = 7,
               formula = y ~ poly(x, 2, raw = T)) +
        
        # Add the R^2
        stat_poly_eq(aes(label = ..rr.label..),
               coef.digits = 2, label.x = 50000, label.y = 1800,
               parse = T, size = 7, formula = y ~ poly(x, 2, raw = T)) +
    
        # Add axes
        ylab(bquote(bold('LiCor PAR (µmol '~m^-2 ~s^-1~')'))) +
        xlab("Sensor Value (Lux)") +
    
        # Add theme
        ggplot.theme

```

Graph showing the predicted values of PAR from the quadratic model graphed against the LiCor

```{r 'Light_predict', echo = F}

#TODO: Add equation to graph

# Construct the graph
light %>% ggplot(aes(x = umol, y = umol.quad)) +
    
        # Add a line with a slope of 1 and intercept of 0
        geom_abline(intercept = 0, slope = 1) + 
    
        # Add points
        geom_point() +
    
        # Add axes
        xlab(bquote(bold('LiCor PAR (µmol '~m^-2 ~s^-1~')'))) +
        ylab(bquote(bold('Predicted Sensor PAR (µmol '~m^-2 ~s^-1~')'))) +
    
        # Add theme
        ggplot.theme

```

This graph compares the BH1750FVI (calibrated to the LiCor) to a [Hobo Microstation](http://www.onsetcomp.com/products/data-loggers/h21-002) with a [S-LIA-M003 PAR Sensor](http://www.onsetcomp.com/products/sensors/s-lia-m003).  The Hobo shows very messy data.  This is the calibration run on the windowsill in TLS in parallel with the soil calibration

```{r 'Light_sensor_comparison', echo = F}

# Construct the plot
light.hobo.comp %>% ggplot(aes(x = umol, y = umol.quad)) +

        # Add points
        geom_point(alpha = 0.4, color = "blue") +
    
        # Add a line with a slope of 1 and intercept of 0
        geom_abline(intercept = 0, slope = 1) + 
    
        # Add axes
        xlab(bquote(bold('Hobo PAR (µmol '~m^-2 ~s^-1~')'))) +
        ylab(bquote(bold('EMU PAR (µmol '~m^-2 ~s^-1~')'))) +
    
        # Title
        ggtitle("Windowsill Calibration") + 
    
        # Add theme
        ggplot.theme

```


### Soil Calibration Results

Calibration of generic silver and gold soil moisture sensor probes with a datalogging scale, measuring water content of the soil every five minutes. The soil used for calibration was a rich loam from the Fenton Meadow. The data columns soil1 and soil2 are gold (ENIG) probes, while soil3 is a silver (HASL) probe.

Concurrently, the soil moisture was measured with a [Hobo Microstation](http://www.onsetcomp.com/products/data-loggers/h21-002) equipped with a [Decagon Devices ECH20 capacitive soil moisture sensor](http://www.decagondevices.eu/products/discontinued-products/ech2o-1).


```{r 'Data_Wrangling_Soil', echo = F, comment = ""}

# From the metadata
fenton.dryweight = 992.7
fenton.wetweight = 1638.7
fenton.density = 0.90

# Data wrangling
cal <- soil %>%

    # Remove the metadata column, and rename soil probes
    select(-metadata, gold1 = soil1, gold2 = soil2, silver = soil3) %>%
    
    # Get rid of any missing data rows
    filter(!is.na(weight)) %>%
    
    # Add columns for measured gravimetric and volumetric water content
    mutate(gwc = (weight - fenton.dryweight) / fenton.dryweight, 
        vwc = gwc * fenton.density) %>%
    
    # Convert the date to date format
    mutate(date = as.POSIXct(date, format = "%m/%d/%y %H:%M"))


```

#### Soil Gold Linear Model

```{r 'Soil_linear_gold_model', echo = F, comment = ""}


# Calibrate the gold probes using a simple linear model
gold.lm <- cal %>%
    
    # Filter for after 7/22, when probes hit a minimum
    filter(date >= as.POSIXct("2017-07-22 00:00")) %>%

    # Gather all the probe measurements into one column
    # This combines data from the two gold probes
    gather(probe, gold, gold1, gold2) %>%
    
    # Run the linear model
    lm(vwc ~ gold, data = .)

# View the model
summary(gold.lm)

```

#### Soil Gold Linear Temperature Model (Combined)

This model is better (by AICc), but it doesn't add much

```{r 'Soil_linear_gold_temp_model', echo = F, comment = ""}


# Calibrate the gold probe with a temperature compensated model
gold.lm.temp <- cal %>%
    
    # Filter for after 7/22, when probes hit a minimum
    filter(date >= as.POSIXct("2017-07-22 00:00")) %>%

    # Gather all the probe measurements into one column
    # This combines data from the two gold probes
    gather(probe, gold, gold1, gold2) %>%
    
    # Run the linear model
    lm(vwc ~ gold + temp, data = .)

# View the model
summary(gold.lm.temp)


# Check AICs, temp-compensated is the best model, but doesn't add much
AICctab(gold.lm, gold.lm.temp)


```

#### Soil Gold Quadratic Model (Combined)

The quadratic model is considerably better

```{r 'Soil_quadratic_gold_model', echo = F, comment = ""}

# Calibrate the gold probes using a quadratic model
gold.quad <- cal %>%
  
    # Filter for after 7/22, when probes hit a minimum
    filter(date >= as.POSIXct("2017-07-22 00:00")) %>%

    # Gather all the probe measurements into one column
    # This combines data from the two gold probes
    gather(probe, gold, gold1, gold2) %>%
    
    # Make a squared column
    mutate(gold.squared = gold ^ 2) %>%

    # Run the quadratic model
    lm(vwc ~ gold + gold.squared, data = .)
 
# View the model
summary(gold.quad)

# Check AICsh
AICctab(gold.lm, gold.lm.temp, gold.quad)

```

#### Soil Gold Quadratic Model (Separate)

Calulates a separate quadratic model for each probe.

I've decided to use the Gold2 probe calibration.  The Gold1 probe did some funny things and was more corroded, so I think it's not so trustworthy.

```{r 'Soil_quadratic_gold_models', echo = F, comment = ""}



# Calibrate the first gold probe using a quadratic model
gold.quad.1 <- cal %>%
  
    # Filter for after 7/15, when probes hit a minimum
    filter(date >= as.POSIXct("2017-07-22 00:00")) %>%
    
    # Make a squared column
    mutate(gold1.squared = gold1 ^ 2) %>%

    # Run the quadratic model
    lm(vwc ~ gold1 + gold1.squared, data = .)
 
# View the model
summary(gold.quad.1)


# Calibrate the first gold probe using a quadratic model
gold.quad.2 <- cal %>%
  
    # Filter for after 7/15, when probes hit a minimum
    filter(date >= as.POSIXct("2017-07-22 00:00")) %>%
    
    # Make a squared column
    mutate(gold2.squared = gold2 ^ 2) %>%

    # Run the quadratic model
    lm(vwc ~ gold2 + gold2.squared, data = .)
 
# View the model
summary(gold.quad.2)


# Rename the intercept coefficient
names(gold.quad.2$coefficients)[1] <- "intercept"

# Write the calibration coefficients to a CSV
write.csv(data.frame(as.list(gold.quad.2$coefficients)), "soil-calibration.csv", 
    row.names = F)


```

#### Soil Silver Linear Model

```{r 'Soil_linear_silver_model', echo = F, comment = ""}


# Calibrate the silver probe using a simple linear model
silver.lm <- cal %>%
    
    # Filter for after 7/22, when probes hit a minimum
    filter(date >= as.POSIXct("2017-07-22 00:00")) %>%
    
    # Run the linear model
    lm(vwc ~ silver, data = .)

# View the model
summary(silver.lm)

```

#### Soil Silver Quadratic Model

```{r 'Soil_quadratic_silver_model', echo = F, comment = ""}


# Calibrate the silver probe using a quadratic model
silver.quad <- cal %>%
    
    # Filter for after 7/22, when probes hit a minimum
    filter(date >= as.POSIXct("2017-07-22 00:00")) %>%
    
    # Make a squared column
    mutate(silver.squared = silver ^ 2) %>%
    
    # Run the linear model
    lm(vwc ~ silver + silver.squared, data = .)

# View the model
summary(silver.quad)

# Check AICc
AICctab(silver.lm, silver.quad)

```



```{r 'Predict_Soil', echo = F, include = F}

# Update the dataset with predicted values from the linear models
cal <- cal %>%
    
    # Add the quadratic calibration for the first gold probe
    mutate(gold1_vwc = gold.quad.1$coefficients[1] + 
        gold.quad.1$coefficients[2] * gold1 + 
        gold.quad.1$coefficients[3] * gold1 ^ 2) %>%
    
    # Add the quadratic calibration for the second gold probe
    mutate(gold2_vwc = gold.quad.2$coefficients[1] + 
        gold.quad.2$coefficients[2] * gold2 + 
        gold.quad.2$coefficients[3] * gold2 ^ 2) %>%
    
    # Add the quadratic calibration for the silver probe
    mutate(silver_vwc = silver.quad$coefficients[1] + 
        silver.quad$coefficients[2] * silver + 
        silver.quad$coefficients[3] * silver ^ 2)


```

#### Soil Calibration Graphs

Showing soil moisture probe values over time. There is some diurnal variation that is likely temperature driven.

Adding temperature to the quadratic fit does improve things slightly, but the temperature range is only ~21-30 C, so it doesn't encompass the full temperature range outside

```{r 'Soil_probes_over_time', echo = F}

cal %>%
    
    # Filter for after 7/22, when probes hit a minimum
    filter(date >= as.POSIXct("2017-07-22 00:00")) %>%
    
    gather(probe, value, gold1, gold2, silver) %>%

    ggplot(aes(x = date, y = value, color = probe, group = probe)) +

        # Set up the colors and legend
        scale_color_manual(values = c("gold", "darkgoldenrod", "gray60"), name = "Probe", 
            labels = c("Gold #1", "Gold #2", "Silver")) +
    
        geom_line(size = 1.2) +
    
        xlab("Time") +
        ylab("Soil Probe Value") + 
    
        # Add theme
        ggplot.theme + 
    
        # Change legend position
        theme(legend.position = c(0.14, 0.85))
    

```

A graph of the calibration and formula using the quadratic model

```{r 'Soil_quad_calibration', echo = F}

cal %>%
    
    # Filter for after 7/22, when probes hit a minimum
    filter(date >= as.POSIXct("2017-07-22 00:00")) %>%

    ggplot(aes(x = gold2, y = vwc)) +
    
        # Add points
        geom_point() + 
    
        # Add linear model
        geom_smooth(method = "lm", color = "red", se = F, 
            formula = y ~ poly(x, 2, raw = T)) + 
    
        # Add the equation
        stat_poly_eq(aes(label = ..eq.label..),
               eq.x.rhs = "italic(x)", coef.digits = 2, label.x = 16600, 
               label.y = 0.30, parse = T, size = 7,
               formula = y ~ poly(x, 2, raw = T)) +
        
        # Add the R^2
        stat_poly_eq(aes(label = ..rr.label..),
               coef.digits = 2, label.x = 16600, label.y = 0.28,
               parse = T, size = 7, formula = y ~ poly(x, 2, raw = T)) +
    
        # Add axis labels
        xlab("Soil Probe Value") + 
        ylab(bquote(bold('Actual VWC'~(m^3/m^3)))) +
    
        # Add theme
        ggplot.theme
    

```

Graph showing the predicted values of VWC from the quadratic model graphed against the actual VWC

```{r 'Soil_predict_comparison', echo = F}

# Plot all the EMU probes vs actual VWC
cal %>%
    
    # Filter for after 7/22, when probes hit a minimum
    filter(date >= as.POSIXct("2017-07-22 00:00")) %>%
    
    # Gather all the probes together
    gather(probe, value, gold1_vwc, gold2_vwc, silver_vwc) %>%
    
    arrange(desc(probe)) %>%

    # Make the plot
    ggplot(aes(x = vwc, y = value, color = probe)) +
    
        # Add a line with a slope of 1 and intercept of 0
        geom_abline(intercept = 0, slope = 1) + 
    
        # Add points
        geom_point() +
    
        # Set up the colors and legend
        scale_color_manual(values = c("gold", "darkgoldenrod", "gray60"), name = "Probe", 
            labels = c("Gold #1", "Gold #2", "Silver")) +
    
        # Add axis labels
        xlab(bquote(bold('Actual VWC'~(m^3/m^3)))) +
        ylab(bquote(bold('Predicted Probe VWC'~(m^3/m^3)))) + 
    
    
        # Add theme
        ggplot.theme + 
    
        # Change legend position
        theme(legend.position = c(0.14, 0.85))
 

```

Comparison of predicted values of VWC from the EMU sensors and VWC from the Hobo Microstation compared to actual VWC

```{r 'Soil_sensor_comparison', echo = F}

# Plot hobo and emu vs actual VWC
cal %>%
    
    # Filter for after 7/22, when probes hit a minimum
    filter(date >= as.POSIXct("2017-07-22 00:00")) %>%
    
    # Gather just the gold2 probe and hobo together
    gather(probe, value, gold2_vwc, hobo_vwc) %>%

    # Make the plot
    ggplot(aes(x = vwc, y = value, color = probe)) +
    
        # Add a line with a slope of 1 and intercept of 0
        geom_abline(intercept = 0, slope = 1) + 
    
        # Add points
        geom_point() +
    
        # Change the graph colors
        scale_color_manual(values = c("black", "blue"), name = "Sensor", 
            labels = c("EMU (calibrated)", "Hobo Microstation")) +
    
        # Add axis labels
        xlab(bquote(bold('Actual VWC'~(m^3/m^3)))) +
        ylab(bquote(bold('Sensor VWC'~(m^3/m^3)))) + 
    
        # Add theme
        ggplot.theme + 
    
        # Change legend position
        theme(legend.position = c(0.22, 0.88))
 

```
